---
title: "Preparing data for DL model fitting"
author: "Scott Forrest"
date: "`r Sys.Date()`"
execute: 
  cache: false
bibliography: references.bib
toc: true
number-sections: false
format: 
  html:
    self-contained: true
    code-fold: show
    code-tools: true
    df-print: paged
    code-line-numbers: true
    code-overflow: scroll
    fig-format: png
    fig-dpi: 300
  pdf:
    geometry: 
      - top=30mm
      - left=30mm
editor:
  source
abstract: |
  This document outlines the steps taken to prepare the data for the deepSSF model fitting.
---

## Loading packages

```{r}
#| warning=FALSE

library(tidyverse)
packages <- c("amt", "sf", "terra", "beepr", "tictoc")
walk(packages, require, character.only = T)

```

## Import data and clean

```{r}

buffalo <- read_csv("data/buffalo.csv")

# remove individuals that have poor data quality or less than about 3 months of data. 
# The "2014.GPS_COMPACT copy.csv" string is a duplicate of ID 2024, so we exlcude it
buffalo <- buffalo %>% filter(!node %in% c("2014.GPS_COMPACT copy.csv", 
                                           # 2005, 2014, 2018, 2021, 2022, 2024,
                                           2029, 2043, 2265, 2284, 2346, 2354))

buffalo <- buffalo %>%  
  group_by(node) %>% 
  arrange(DateTime, .by_group = T) %>% 
  distinct(DateTime, .keep_all = T) %>% 
  arrange(node) %>% 
  mutate(ID = node)

buffalo_clean <- buffalo[, c(12, 2, 4, 3)]
colnames(buffalo_clean) <- c("id", "time", "lon", "lat")
attr(buffalo_clean$time, "tzone") <- "Australia/Queensland"
head(buffalo_clean)
tz(buffalo_clean$time)

buffalo_ids <- unique(buffalo_clean$id)

```

## Setup trajectory

Use the `amt` package to create a trajectory object from the cleaned data. 

```{r}

buffalo_all <- buffalo_clean %>% mk_track(id = id,
                                           lon,
                                           lat, 
                                           time, 
                                           all_cols = T,
                                           crs = 4326) %>% 
  transform_coords(crs_to = 3112, crs_from = 4326) # Transformation to GDA94 / 
# Geoscience Australia Lambert (https://epsg.io/3112)

```

## Plot the data coloured by time

```{r}

buffalo_all %>%
  ggplot(aes(x = x_, y = y_, colour = t_)) +
  geom_point(alpha = 0.25, size = 1) + 
  coord_fixed() +
  scale_colour_viridis_c() +
  theme_classic()

```

## Reading in the environmental covariates

```{r}

ndvi_projected <- rast("mapping/cropped rasters/ndvi_GEE_projected_watermask20230207.tif")
terra::time(ndvi_projected) <- as.POSIXct(lubridate::ymd("2018-01-01") + months(0:23))
slope <- rast("mapping/cropped rasters/slope_raster.tif")
veg_herby <- rast("mapping/cropped rasters/veg_herby.tif")
canopy_cover <- rast("mapping/cropped rasters/canopy_cover.tif")

# change the names (these will become the column names when extracting 
# covariate values at the used and random steps)
names(ndvi_projected) <- rep("ndvi", terra::nlyr(ndvi_projected))
names(slope) <- "slope"
names(veg_herby) <- "veg_herby"
names(canopy_cover) <- "canopy_cover"

# to plot the rasters
plot(ndvi_projected)
plot(slope)
plot(veg_herby)
plot(canopy_cover)

```

## Sentinel-2 spectral layers

Create vector of dates. We just use the 15th of the month (the layers were averaged across the month). 

```{r}

dates <- as.POSIXct(lubridate::ymd("2019-01-15")) + months(0:11)

```

### Import the layers and add times

```{r}

### January 2019
s2_10m_2019_01 <- rast("mapping/cropped rasters/sentinel2/10m/S2_SR_masked_2019_01.tif")
terra::time(s2_10m_2019_01) <- rep(as.POSIXct(lubridate::ymd("2019-01-15"), tz = "Australia/Queensland"), nlyr(s2_10m_2019_01))
s2_25m_2019_01 <- terra::resample(s2_10m_2019_01, ndvi_projected, method = "bilinear")
s2_25m_2019_01 <- s2_25m_2019_01 / 10000

### February 2019
s2_10m_2019_02 <- rast("mapping/cropped rasters/sentinel2/10m/S2_SR_masked_2019_02.tif")
terra::time(s2_10m_2019_02) <- rep(as.POSIXct(lubridate::ymd("2019-02-15"), tz = "Australia/Queensland"), nlyr(s2_10m_2019_02))
s2_25m_2019_02 <- terra::resample(s2_10m_2019_02, ndvi_projected, method = "bilinear")
s2_25m_2019_02 <- s2_25m_2019_02 / 10000

### March 2019
s2_10m_2019_03 <- rast("mapping/cropped rasters/sentinel2/10m/S2_SR_masked_2019_03.tif")
terra::time(s2_10m_2019_03) <- rep(as.POSIXct(lubridate::ymd("2019-03-15"), tz = "Australia/Queensland"), nlyr(s2_10m_2019_03))
s2_25m_2019_03 <- terra::resample(s2_10m_2019_03, ndvi_projected, method = "bilinear")
s2_25m_2019_03 <- s2_25m_2019_03 / 10000

### April 2019
s2_10m_2019_04 <- rast("mapping/cropped rasters/sentinel2/10m/S2_SR_masked_2019_04.tif")
terra::time(s2_10m_2019_04) <- rep(as.POSIXct(lubridate::ymd("2019-04-15"), tz = "Australia/Queensland"), nlyr(s2_10m_2019_04))
s2_25m_2019_04 <- terra::resample(s2_10m_2019_04, ndvi_projected, method = "bilinear")
s2_25m_2019_04 <- s2_25m_2019_04 / 10000

### May 2019
s2_10m_2019_05 <- rast("mapping/cropped rasters/sentinel2/10m/S2_SR_masked_2019_05.tif")
terra::time(s2_10m_2019_05) <- rep(as.POSIXct(lubridate::ymd("2019-05-15"), tz = "Australia/Queensland"), nlyr(s2_10m_2019_05))
s2_25m_2019_05 <- terra::resample(s2_10m_2019_05, ndvi_projected, method = "bilinear")
s2_25m_2019_05 <- s2_25m_2019_05 / 10000

### June 2019
s2_10m_2019_06 <- rast("mapping/cropped rasters/sentinel2/10m/S2_SR_masked_2019_06.tif")
# plot(s2_10m_2019_06[[2]])
terra::time(s2_10m_2019_06) <- rep(as.POSIXct(lubridate::ymd("2019-06-15"), tz = "Australia/Queensland"), nlyr(s2_10m_2019_06))
s2_25m_2019_06 <- terra::resample(s2_10m_2019_06, ndvi_projected, method = "bilinear")
s2_25m_2019_06 <- s2_25m_2019_06 / 10000

### July 2019
s2_10m_2019_07 <- rast("mapping/cropped rasters/sentinel2/10m/S2_SR_masked_2019_07.tif")
terra::time(s2_10m_2019_07) <- rep(as.POSIXct(lubridate::ymd("2019-07-15"), tz = "Australia/Queensland"), nlyr(s2_10m_2019_07))
s2_25m_2019_07 <- terra::resample(s2_10m_2019_07, ndvi_projected, method = "bilinear")
s2_25m_2019_07 <- s2_25m_2019_07 / 10000

### August 2019
s2_10m_2019_08 <- rast("mapping/cropped rasters/sentinel2/10m/S2_SR_masked_2019_08.tif")
terra::time(s2_10m_2019_08) <- rep(as.POSIXct(lubridate::ymd("2019-08-15"), tz = "Australia/Queensland"), nlyr(s2_10m_2019_08))
s2_25m_2019_08 <- terra::resample(s2_10m_2019_08, ndvi_projected, method = "bilinear")
s2_25m_2019_08 <- s2_25m_2019_08 / 10000

### September 2019
s2_10m_2019_09 <- rast("mapping/cropped rasters/sentinel2/10m/S2_SR_masked_2019_09.tif")
terra::time(s2_10m_2019_09) <- rep(as.POSIXct(lubridate::ymd("2019-09-15"), tz = "Australia/Queensland"), nlyr(s2_10m_2019_09))
s2_25m_2019_09 <- terra::resample(s2_10m_2019_09, ndvi_projected, method = "bilinear")
s2_25m_2019_09 <- s2_25m_2019_09 / 10000

### October 2019
s2_10m_2019_10 <- rast("mapping/cropped rasters/sentinel2/10m/S2_SR_masked_2019_10.tif")
terra::time(s2_10m_2019_10) <- rep(as.POSIXct(lubridate::ymd("2019-10-15"), tz = "Australia/Queensland"), nlyr(s2_10m_2019_10))
s2_25m_2019_10 <- terra::resample(s2_10m_2019_10, ndvi_projected, method = "bilinear")
s2_25m_2019_10 <- s2_25m_2019_10 / 10000

### November 2019
s2_10m_2019_11 <- rast("mapping/cropped rasters/sentinel2/10m/S2_SR_masked_2019_11.tif")
terra::time(s2_10m_2019_11) <- rep(as.POSIXct(lubridate::ymd("2019-11-15"), tz = "Australia/Queensland"), nlyr(s2_10m_2019_11))
s2_25m_2019_11 <- terra::resample(s2_10m_2019_11, ndvi_projected, method = "bilinear")
s2_25m_2019_11 <- s2_25m_2019_11 / 10000

### December 2019
s2_10m_2019_12 <- rast("mapping/cropped rasters/sentinel2/10m/S2_SR_masked_2019_12.tif")
terra::time(s2_10m_2019_12) <- rep(as.POSIXct(lubridate::ymd("2019-12-15"), tz = "Australia/Queensland"), nlyr(s2_10m_2019_12))
s2_25m_2019_12 <- terra::resample(s2_10m_2019_12, ndvi_projected, method = "bilinear")
s2_25m_2019_12 <- s2_25m_2019_12 / 10000

```

### Combine into a list

```{r}

s2_list <- list(s2_25m_2019_01, s2_25m_2019_02, s2_25m_2019_03, 
                s2_25m_2019_04, s2_25m_2019_05, s2_25m_2019_06, 
                s2_25m_2019_07, s2_25m_2019_08, s2_25m_2019_09, 
                s2_25m_2019_10, s2_25m_2019_11, s2_25m_2019_12)

names(s2_25m_2019_12)

```

# Generating the data to fit a deepSSF model

## Set up the spatial extent of the local covariates

```{r}

buffalo_all <- buffalo_all %>%  
  arrange(id, t_)

# create a vector of ids
buffalo_ids <- unique(buffalo_all$id)

# get the resolution from the covariates
res <- terra::res(ndvi_projected)[1]

# how much to trim on either side of the location, 
# this will determine the extent of the spatial inputs to the deepSSF model
buffer <- 1250 + (res/2)
# calculate the number of cells in each axis
nxn_cells <- buffer*2/res

# hourly lag - to set larger time differences between locations
hourly_lag <- 1

```

Loop over each individual and save the local rasters

```{r}

for(i in 1:length(buffalo_ids)) {

  buffalo_data <- buffalo_all %>% filter(id == buffalo_ids[i])
  # all data for that individual
  # buffalo_data <- buffalo_data %>% arrange(t_)
  # a subset of the data for that individual - mostly for testing
  buffalo_data <- buffalo_data %>% arrange(t_) %>% slice(1:10)
  
  n_samples <- nrow(buffalo_data)
  
  tic()
  
  buffalo_data_covs <- buffalo_data %>% mutate(
    
    x1_ = x_,
    y1_ = y_,
    x2_ = lead(x1_, n = hourly_lag, default = NA),
    y2_ = lead(y1_, n = hourly_lag, default = NA),
    x2_cent = x2_ - x1_,
    y2_cent = y2_ - y1_,
    t2_ = lead(t_, n = hourly_lag, default = NA),
    t_diff = round(difftime(t2_, t_, units = "hours"),0),
    hour_t1 = lubridate::hour(t_),
    yday_t1 = lubridate::yday(t_),
    hour_t2 = lubridate::hour(t2_),
    hour_t2_sin = sin(2*pi*hour_t2/24),
    hour_t2_cos = cos(2*pi*hour_t2/24),
    yday_t2 = lubridate::yday(t2_),
    yday_t2_sin = sin(2*pi*yday_t2/365.25),
    yday_t2_cos = cos(2*pi*yday_t2/365.25),
    
    sl = c(sqrt(diff(y_)^2 + diff(x_)^2), NA),
    log_sl = log(sl),
    bearing = c(atan2(diff(y_), diff(x_)), NA),
    bearing_sin = sin(bearing),
    bearing_cos = cos(bearing),
    ta = c(NA, ifelse(
      diff(bearing) > pi, diff(bearing)-(2*pi), ifelse(
        diff(bearing) < -pi, diff(bearing)+(2*pi), diff(bearing)))),
    cos_ta = cos(ta),
      
    # extent for cropping the spatial covariates
    x_min = x_ - buffer,
    x_max = x_ + buffer,
    y_min = y_ - buffer,
    y_max = y_ + buffer
    
  
    # crop out and store the local covariates centered on the animal's location 
    # with an extent set in the previous chunk
  ) %>% rowwise() %>% mutate(

    extent_00centre = list(ext(x_min - x_, x_max - x_, y_min - y_, y_max - y_)),
    
    # Sentinel 2
    s2_index = which.min(abs(lubridate::month(t_) - lubridate::month(dates))),
    
    # Band 1
    s2_b1_cent = list({
      s2_b1_cent = crop( s2_list[[s2_index]][[1]], ext(x_min, x_max, y_min, y_max))
      ext(s2_b1_cent) <- extent_00centre
      s2_b1_cent
      }),
    
    # Band 2
    s2_b2_cent = list({
      s2_b2_cent = crop( s2_list[[s2_index]][[2]], ext(x_min, x_max, y_min, y_max))
      ext(s2_b2_cent) <- extent_00centre
      s2_b2_cent
      }),
    
    # Band 3
    s2_b3_cent = list({
      s2_b3_cent = crop( s2_list[[s2_index]][[3]], ext(x_min, x_max, y_min, y_max))
      ext(s2_b3_cent) <- extent_00centre
      s2_b3_cent
      }),
    
    # Band 4
    s2_b4_cent = list({
      s2_b4_cent = crop( s2_list[[s2_index]][[4]], ext(x_min, x_max, y_min, y_max))
      ext(s2_b4_cent) <- extent_00centre
      s2_b4_cent
      }),
    
    # Band 5
    s2_b5_cent = list({
      s2_b5_cent = crop( s2_list[[s2_index]][[5]], ext(x_min, x_max, y_min, y_max))
      ext(s2_b5_cent) <- extent_00centre
      s2_b5_cent
      }),
    
    # Band 6
    s2_b6_cent = list({
      s2_b6_cent = crop( s2_list[[s2_index]][[6]], ext(x_min, x_max, y_min, y_max))
      ext(s2_b6_cent) <- extent_00centre
      s2_b6_cent
      }),
    
    # Band 7
    s2_b7_cent = list({
      s2_b7_cent = crop( s2_list[[s2_index]][[7]], ext(x_min, x_max, y_min, y_max))
      ext(s2_b7_cent) <- extent_00centre
      s2_b7_cent
      }),
    
    # Band 8
    s2_b8_cent = list({
      s2_b8_cent = crop( s2_list[[s2_index]][[8]], ext(x_min, x_max, y_min, y_max))
      ext(s2_b8_cent) <- extent_00centre
      s2_b8_cent
      }),
    
    # Band 8A
    s2_b8a_cent = list({
      s2_b8a_cent = crop( s2_list[[s2_index]][[9]], ext(x_min, x_max, y_min, y_max))
      ext(s2_b8a_cent) <- extent_00centre
      s2_b8a_cent
      }),
    
    # Band 9
    s2_b9_cent = list({
      s2_b9_cent = crop( s2_list[[s2_index]][[10]], ext(x_min, x_max, y_min, y_max))
      ext(s2_b9_cent) <- extent_00centre
      s2_b9_cent
      }),
    
    # Band 11
    s2_b11_cent = list({
      s2_b11_cent = crop( s2_list[[s2_index]][[11]], ext(x_min, x_max, y_min, y_max))
      ext(s2_b11_cent) <- extent_00centre
      s2_b11_cent
      }),
    
    # Band 12
    s2_b12_cent = list({
      s2_b12_cent = crop( s2_list[[s2_index]][[12]], ext(x_min, x_max, y_min, y_max))
      ext(s2_b12_cent) <- extent_00centre
      s2_b12_cent
      }),

    # NDVI
    ndvi_index = which.min(abs(difftime(t_, terra::time(ndvi_projected)))),
    ndvi_cent = list({
      ndvi_cent = crop(ndvi_projected[[ndvi_index]], ext(x_min, x_max, y_min, y_max))
      ext(ndvi_cent) <- extent_00centre
      ndvi_cent
      }),

    # herbaceous vegetation
    veg_herby_cent = list({
      veg_herby_cent = crop(veg_herby, ext(x_min, x_max, y_min, y_max))
      ext(veg_herby_cent) <- extent_00centre
      veg_herby_cent
      }),

    # canopy cover
    canopy_cover_cent = list({
      canopy_cover_cent = crop(canopy_cover, ext(x_min, x_max, y_min, y_max))
      ext(canopy_cover_cent) <- extent_00centre
      canopy_cover_cent
      }),

    # slope

    slope_cent = list({
      slope_cent <- crop(slope, ext(x_min, x_max, y_min, y_max))
      ext(slope_cent) <- extent_00centre
      slope_cent
      }),

    # rasterised location of the next step - centred on (0,0)
    points_vect_cent = list(terra::vect(cbind(x2_ - x_, y2_ - y_), type = "points", crs = "EPSG:3112")),
    pres_cent = list(rasterize(points_vect_cent, s2_b1_cent, background=0))

  ) %>% ungroup() # to remove the 'rowwise' class


  toc()

  # remove steps that fall outside of the local spatial extent
  buffalo_data_covs <- buffalo_data_covs %>%
    filter(x2_cent > -buffer & x2_cent < buffer & y2_cent > -buffer & y2_cent < buffer) %>%
    drop_na(ta)


  buffalo_data_df <- buffalo_data_covs %>%
  dplyr::select(-extent_00centre,
                -s2_b1_cent,
                -s2_b2_cent,
                -s2_b3_cent,
                -s2_b4_cent,
                -s2_b5_cent,
                -s2_b6_cent,
                -s2_b7_cent,
                -s2_b8_cent,
                -s2_b8a_cent,
                -s2_b9_cent,
                -s2_b11_cent,
                -s2_b12_cent,
                -ndvi_cent,
                -veg_herby_cent,
                -canopy_cover_cent,
                -slope_cent,
                -points_vect_cent,
                -pres_cent
                )

  # save the data
  write_csv(buffalo_data_df, paste0("buffalo_local_data_id/buffalo_", buffalo_ids[i],
                                    "_data_df_lag_", hourly_lag, "hr_n", n_samples, ".csv"))

  # saving the raster objects
  
  rast(buffalo_data_covs$s2_b1_cent) %>%
    writeRaster(paste0("buffalo_local_layers_id/buffalo_", buffalo_ids[i], "_s2_b1_cent",
                       nxn_cells, "x", nxn_cells, "_lag_", hourly_lag, "hr_n", n_samples, ".tif"),
                overwrite = T)
  
  rast(buffalo_data_covs$s2_b2_cent) %>%
    writeRaster(paste0("buffalo_local_layers_id/buffalo_", buffalo_ids[i], "_s2_b2_cent",
                       nxn_cells, "x", nxn_cells, "_lag_", hourly_lag, "hr_n", n_samples, ".tif"),
                overwrite = T)
  
  rast(buffalo_data_covs$s2_b3_cent) %>% 
    writeRaster(paste0("buffalo_local_layers_id/buffalo_", buffalo_ids[i], "_s2_b3_cent",
                       nxn_cells, "x", nxn_cells, "_lag_", hourly_lag, "hr_n", n_samples, ".tif"),
                overwrite = T)
  
  rast(buffalo_data_covs$s2_b4_cent) %>%
    writeRaster(paste0("buffalo_local_layers_id/buffalo_", buffalo_ids[i], "_s2_b4_cent",
                       nxn_cells, "x", nxn_cells, "_lag_", hourly_lag, "hr_n", n_samples, ".tif"),
                overwrite = T)
  
  rast(buffalo_data_covs$s2_b5_cent) %>%
    writeRaster(paste0("buffalo_local_layers_id/buffalo_", buffalo_ids[i], "_s2_b5_cent",
                       nxn_cells, "x", nxn_cells, "_lag_", hourly_lag, "hr_n", n_samples, ".tif"),
                overwrite = T)
  
  rast(buffalo_data_covs$s2_b6_cent) %>%
    writeRaster(paste0("buffalo_local_layers_id/buffalo_", buffalo_ids[i], "_s2_b6_cent",
                       nxn_cells, "x", nxn_cells, "_lag_", hourly_lag, "hr_n", n_samples, ".tif"),
                overwrite = T)
  
  rast(buffalo_data_covs$s2_b7_cent) %>%
    writeRaster(paste0("buffalo_local_layers_id/buffalo_", buffalo_ids[i], "_s2_b7_cent",
                       nxn_cells, "x", nxn_cells, "_lag_", hourly_lag, "hr_n", n_samples, ".tif"),
                overwrite = T)
  
  rast(buffalo_data_covs$s2_b8_cent) %>%
    writeRaster(paste0("buffalo_local_layers_id/buffalo_", buffalo_ids[i], "_s2_b8_cent",
                       nxn_cells, "x", nxn_cells, "_lag_", hourly_lag, "hr_n", n_samples, ".tif"),
                overwrite = T)
  
  rast(buffalo_data_covs$s2_b8a_cent) %>%
    writeRaster(paste0("buffalo_local_layers_id/buffalo_", buffalo_ids[i], "_s2_b8a_cent",
                       nxn_cells, "x", nxn_cells, "_lag_", hourly_lag, "hr_n", n_samples, ".tif"),
                overwrite = T)
  
  rast(buffalo_data_covs$s2_b9_cent) %>%
    writeRaster(paste0("buffalo_local_layers_id/buffalo_", buffalo_ids[i], "_s2_b9_cent",
                       nxn_cells, "x", nxn_cells, "_lag_", hourly_lag, "hr_n", n_samples, ".tif"),
                overwrite = T)
  
  rast(buffalo_data_covs$s2_b11_cent) %>%
    writeRaster(paste0("buffalo_local_layers_id/buffalo_", buffalo_ids[i], "_s2_b11_cent",
                       nxn_cells, "x", nxn_cells, "_lag_", hourly_lag, "hr_n", n_samples, ".tif"),
                overwrite = T)
  
  rast(buffalo_data_covs$s2_b12_cent) %>%
    writeRaster(paste0("buffalo_local_layers_id/buffalo_", buffalo_ids[i], "_s2_b12_cent",
                       nxn_cells, "x", nxn_cells, "_lag_", hourly_lag, "hr_n", n_samples, ".tif"),
                overwrite = T)
  
  rast(buffalo_data_covs$pres_cent) %>%
    writeRaster(paste0("buffalo_local_layers_id/buffalo_", buffalo_ids[i], "_pres_cent",
                       nxn_cells, "x", nxn_cells, "_lag_", hourly_lag, "hr_n", n_samples, ".tif"),
                overwrite = T)
  
}

```

## Check the outputs

```{r}

head(buffalo_data_covs)
which(is.na(buffalo_data_covs$ta))

```

## Plot a subset of the local covariates

```{r}

n_plots <- 5

# Blue layer
walk(buffalo_data_covs$s2_b2_cent[1:n_plots], terra::plot)
# Red layer
walk(buffalo_data_covs$s2_b3_cent[1:n_plots], terra::plot)
# Green layer
walk(buffalo_data_covs$s2_b4_cent[1:n_plots], terra::plot)

# Plot as RGB image
rgb_image <- c(buffalo_data_covs$s2_b4_cent[[5]]*1e4, buffalo_data_covs$s2_b3_cent[[5]]*1e4, buffalo_data_covs$s2_b2_cent[[5]]*1e4)

# Save plot as a PNG image
png("outputs/rgb_local_env.png", width = 120, height = 120, units = "mm", res = 600)  # Adjust width, height, and resolution as needed
terra::plotRGB(rgb_image, r = 1, g = 2, b = 3, 
               smooth = FALSE, mar = 1.5, axes = TRUE)  # adjust 'scale' if needed
dev.off()  # Close the graphic device to save the file

# find which coordinates the 1 is
buffalo_data_covs$pres_cent[1]
which(buffalo_data_covs$pres_cent[1] == 1)

# Find the cell index of the first occurrence of 1
cell_index <- which(values(buffalo_data_covs$pres_cent[[5]]) == 1)[1]  # First occurrence

# Get the coordinates of this cell
coordinates <- terra::xyFromCell(buffalo_data_covs$pres_cent[[5]], cell_index)
buffalo_data_covs$s2_b2_cent[1]

```

