---
title: "deepSSF Data Prep"
author: "Scott Forrest"
date: "`r Sys.Date()`"
execute: 
  cache: false
bibliography: references.bib
toc: true
number-sections: false
format: 
  html:
    # self-contained: true
    code-fold: show
    code-tools: true
    df-print: paged
    code-line-numbers: true
    code-overflow: scroll
    fig-format: png
    fig-dpi: 300
  pdf:
    geometry: 
      - top=30mm
      - left=30mm
editor:
  source
abstract: |
  Here we prepare the data for fitting a deepSSF model. We load the GPS data and tidy it, create a trajectory object containing a series of steps and read in the environmental layers. For each observed step we crop out local subsets of the environmental layers, centred on the step's location. For every step we will then have the surrounding environmental information stored as a stack of small rasters, with the 'target' of what we're trying to predict (the actual location of the next step) as a raster layer as well, with all cells being 0 except the observed next location, which is 1. These, along with temporal covariates and the previous bearing, will be the input data for the deepSSF model.
---

## Loading packages

```{r}
#| warning=FALSE

library(tidyverse)
packages <- c("amt", "sf", "terra", "beepr", "tictoc")
walk(packages, require, character.only = T)

```

## Import data

```{r}

buffalo <- read_csv("data/SpaceCows/Mimal_processed.csv")

buffalo <- buffalo %>% mutate(
  datetime = lubridate::ymd_hms(paste(year, month, day, time), tz = "UTC"),
  .before = TS
)

head(buffalo)

buffalo$TS

```

## Tidy data

```{r}

# arrange by time and exclude any duplicate timestamps (within the same individual's data)
buffalo <- buffalo %>%  
  group_by(device_id) %>% 
  arrange(TS, .by_group = T) %>% 
  # distinct(DateTime, .keep_all = T) %>% 
  arrange(device_id) %>% 
  mutate(id = device_id, 
         datetime = TS,
         .after = TS)

# keep only the relevant columns
buffalo_clean <- buffalo[, c("id", "datetime", "Longitude", "Latitude")]

# rename the columns
colnames(buffalo_clean) <- c("id", "time", "lon", "lat")

# convert the time to the correct timezone
attr(buffalo_clean$time, "tzone") <- "Australia/Queensland"
head(buffalo_clean)

# check timezone
tz(buffalo_clean$time)

# create a vector of the unique ids for indexing later
buffalo_ids <- unique(buffalo_clean$id)

min(buffalo_clean$lon)
max(buffalo_clean$lon)
min(buffalo_clean$lat)
max(buffalo_clean$lat)

```

## Setup trajectory

Use the `amt` package to create a trajectory object from the cleaned data. 

```{r}

buffalo_all <- buffalo_clean %>% mk_track(id = id,
                                          lon,
                                          lat, 
                                          time, 
                                          all_cols = T,
                                          crs = 4326) %>% 
  # Transformation to GDA94 / Geoscience Australia Lambert (https://epsg.io/3112)
  transform_coords(crs_to = 3112, crs_from = 4326) %>% arrange(id) 

```

## Plot the data coloured by time

```{r}

buffalo_all %>%
  ggplot(aes(x = x_, y = y_, colour = id)) +
  geom_point(alpha = 0.5, size = 0.1) + 
  coord_fixed() +
  scale_x_continuous("Easting (m)") +
  scale_y_continuous("Northing (m)") +
  scale_colour_viridis_d() +
  theme_classic() +
  theme(legend.position = "none") 

```

There are a few locations to the north of the Bigedi cluster (bottom right) that stray a fair way from the main cluster. Normally we wouldn't worry much about these locations, but we want to set the extent of the landscape based on the locations, and including these few locations means we would have to run the simulations over a much larger area than we need to, without adding much information from including the datapoints. We will therefore remove the locations from these two individuals, which only have 280 and 51 GPS fixes.

```{r}

# identify which IDs are to the top right of the Bigedi cluster
buffalo_all %>% filter(x_ > 6e4 & y_ > -1.58e6)

buffalo_all %>% filter(id == "009503764")
buffalo_all %>% filter(id == "017250804")

buffalo_all <- buffalo_all %>% filter(!id %in% c("009503764", "017250804"))

```

## Check the plot again

```{r}

buffalo_all %>% 
  ggplot(aes(x = x_, y = y_, colour = id)) +
  geom_point(alpha = 0.5, size = 0.1) + 
  coord_fixed() +
  scale_x_continuous("Easting (m)") +
  scale_y_continuous("Northing (m)") +
  scale_colour_viridis_d() +
  theme_classic() +
  theme(legend.position = "none") 

ggsave("outputs/data_prep/buffalo_mimal_map.png", 
       width = 150, height = 150, units = "mm",  dpi = 1000)

# colour by index to see the order of the points
buffalo_all %>% mutate(index = 1:n()) %>%
  ggplot(aes(x = x_, y = y_, colour = index)) +
  geom_point(alpha = 0.5, size = 0.1) + 
  coord_fixed() +
  scale_x_continuous("Easting (m)") +
  scale_y_continuous("Northing (m)") +
  scale_colour_viridis_c() +
  theme_classic() +
  theme(legend.position = "right") 

```

## Cluster into two groups using k-means

```{r}

# set seed
set.seed(42)

# k-means clustering to group the data into two groups
buffalo_all_kmeans <- buffalo_all %>%
  dplyr::select(x_, y_) %>%
  kmeans(centers = 2, nstart = 25)

# add the cluster to the data
buffalo_all <- buffalo_all %>%
  mutate(cluster = as.factor(buffalo_all_kmeans$cluster))

# number of IDs in each cluster
buffalo_all %>%
  group_by(cluster) %>%
  summarise(n_ids = n_distinct(id),
            n_locations = n())

```

## Plot the data coloured by cluster

```{r}

buffalo_all %>%
  ggplot(aes(x = x_, y = y_, colour = cluster)) +
  geom_point(alpha = 0.5, size = 0.1) + 
  coord_fixed() +
  scale_x_continuous("Easting (m)") +
  scale_y_continuous("Northing (m)") +
  scale_colour_viridis_d() +
  theme_classic() +
  theme(legend.position = "bottom") 

ggsave("outputs/data_prep/buffalo_mimal_cluster_map.png", 
       width = 150, height = 150, units = "mm",  dpi = 1000)

```

## Plot a timeline of each individual

```{r}

buffalo_all %>% 
  ggplot(aes(x = t_, y = as.factor(id), colour = as.factor(id))) +
  geom_point(alpha = 0.5, size = 0.01) + 
  scale_x_datetime("Date") +
  scale_y_discrete("Individual ID") +
  scale_colour_viridis_d() +
  theme_classic() +
  theme(legend.position = "none",
        axis.text.y = element_blank()) 

ggsave("outputs/data_prep/buffalo_mimal_timeline.png", 
       width = 150, height = 150, units = "mm",  dpi = 600)

buffalo_all %>% 
  ggplot(aes(x = t_, y = as.factor(id), colour = cluster)) +
  geom_point(alpha = 0.5, size = 0.01) + 
  scale_x_datetime("Date") +
  scale_y_discrete("Individual ID") +
  scale_colour_viridis_d() +
  theme_classic() +
  theme(legend.position = "bottom",
        axis.text.y = element_blank()) 

ggsave("outputs/data_prep/buffalo_mimal_cluster_timeline.png", 
       width = 150, height = 150, units = "mm",  dpi = 600)

```

## Number of location per day

```{r}

buffalo_daily_locations <- buffalo_all %>% 
  mutate(date = lubridate::date(t_)) %>% 
  dplyr::group_by(date, cluster) %>% 
  summarise(
    n_locations = n(),
    n_individuals = n_distinct(id)
)

buffalo_daily_locations %>% 
  ggplot() + 
    geom_col(aes(x = date, y = n_locations), 
             fill = "skyblue") + 
    geom_col(aes(x = date, y = n_individuals), 
             fill = "orange") + 
    scale_x_date("Date") +
    scale_y_continuous("Number of IDs/Locations") +
    scale_colour_viridis_d() +
    theme_classic() +
    theme(legend.position = "none") 

ggsave("outputs/data_prep/buffalo_mimal_daily_locations.png",
       width = 150, height = 80, units = "mm",  dpi = 600)

```

## Step lengths at the different sites

```{r}

buffalo_daily_locations %>% 
  # filter(date <= "2023-12-31") %>% 
  ggplot() + 
    geom_col(aes(x = date, y = n_individuals, fill = cluster),
             alpha = 1,
             position = "dodge") +
    scale_x_date("Date") +
    scale_y_continuous("Number of IDs/Locations") +
    scale_colour_viridis_d() +
    theme_classic() +
    theme(legend.position = "bottom") 

```

## Number of locations per individual

```{r}

buffalo_id_locations <- buffalo_all %>% 
  dplyr::group_by(id, cluster) %>% 
  summarise(
    n_locations = n()
)

buffalo_id_locations %>% 
  ggplot() + 
    geom_histogram(aes(x = n_locations, 
             fill = cluster),
             alpha = 0.5,
             binwidth = 20) + 
    scale_x_continuous("Locations per ID") +
    scale_y_continuous("Number of IDs") +
    scale_colour_viridis_d() +
    theme_classic() +
    theme(legend.position = "none") 

ggsave("outputs/data_prep/buffalo_mimal_daily_locations.png",
       width = 150, height = 80, units = "mm",  dpi = 600)

```


# Generating the data to fit a deepSSF model

## Create a steps object to check the step lengths

```{r}

# nest the data by individual
buffalo_all_nested <- buffalo_all %>% arrange(id) %>% nest(data = -"id")

buffalo_all_nested_steps <- buffalo_all_nested %>%
  mutate(steps = map(data, function(x)
    x %>% 
      steps(keep_cols = 'start')))

# unnest the data after creating 'steps' objects
buffalo_all_steps <- buffalo_all_nested_steps %>% 
  amt::select(id, steps) %>% 
  amt::unnest(cols = steps)

head(buffalo_all_steps)

```

## Prepare and export data for single layer saving (Python script)

We also have another script that saves all local layers as individual npy arrays, which works better when there are many steps per id (to prevent large tif files), and can make things a bit easier for training models with many ids.

```{r}

buffalo_all_steps <- buffalo_all_steps %>% mutate(
  
  # difference in the x and y directions
  dx = x2_ - x1_,
  dy = y2_ - y1_,
  
  # time difference between the two locations
  dt_minute = as.numeric(round(difftime(t2_, t1_, units = "mins"), 0)),
  dt_hour = as.numeric(round(difftime(t2_, t1_, units = "hours"), 2)),
  
  # add the minutes as a decimal to the hour, and the hour as a decimal to the day
  hour_t1 = round(lubridate::hour(t1_) + lubridate::minute(t1_)/60,2),
  yday_t1 = round(lubridate::yday(t1_) + lubridate::hour(t1_)/24,2),
  
  hour_t1_sin1 = sin(2*pi*hour_t1/24),
  hour_t1_cos1 = cos(2*pi*hour_t1/24),
  hour_t1_sin2 = sin(4*pi*hour_t1/24),
  hour_t1_cos2 = cos(4*pi*hour_t1/24),
  
  yday_t1_sin1 = sin(2*pi*yday_t1/365.25),
  yday_t1_cos1 = cos(2*pi*yday_t1/365.25),
  yday_t1_sin2 = sin(4*pi*yday_t1/365.25),
  yday_t1_cos2 = cos(4*pi*yday_t1/365.25),
  
  # add the minutes as a decimal to the hour, and the hour as a decimal to the day
  hour_t2 = round(lubridate::hour(t2_) + lubridate::minute(t2_)/60,2),
  yday_t2 = round(lubridate::yday(t2_) + lubridate::hour(t2_)/24,2),
  
  bearing = direction_p,
  bearing_sin = sin(bearing),
  bearing_cos = cos(bearing)
  
)

buffalo_all_steps

```

### Plot step lengths

```{r}

buffalo_all_steps %>% 
  filter(dt_hour < 24) %>%
  filter(sl_ < 5000) %>%
  ggplot() +
  geom_density(aes(x = sl_), #, bins = 50
                 fill = "skyblue", colour = "skyblue", alpha = 0.75) +
  scale_x_continuous("Step length (m)") + #, limits = c(-25, 1250)
  scale_x_log10("Step length (m)") + #, limits = c(-25, 1250)
  scale_y_continuous("Density") +
  ggtitle("Step lengths") +
  theme_classic()

```

### Plot step lengths for each hour

```{r}

for(i in 0:23) {
  
  print(buffalo_all_steps %>% 
    filter(lubridate::hour(t1_) == i) %>%
    filter(sl_ < 5000) %>%
    ggplot() +
    geom_density(aes(x = sl_), #, bins = 50
                   fill = "skyblue", colour = "skyblue", alpha = 0.75) +
    scale_x_continuous("Step length (m)") + #, limits = c(-25, 1250)
    # scale_x_log10("Step length (m)") + #, limits = c(-25, 1250)
    scale_y_continuous("Density") +
    ggtitle(paste("Step lengths at hour", i)) +
    theme_classic())
  
}

```

### Step lengths vs step duration

```{r}

buffalo_all_steps %>% 
  filter(dt_hour < 24) %>%
  ggplot() +
  geom_point(aes(x = dt_hour, y = sl_), 
                 fill = "black", colour = "black", 
             alpha = 0.05, size = 0.1) +
  scale_x_continuous("Time between GPS locations (hours)") + #, limits = c(-25, 1250)
  scale_y_continuous("Step length (m)") +
  # scale_x_log10("Time between GPS locations (hours)") + #, limits = c(-25, 1250)
  scale_y_log10("Step length (m)") +
  ggtitle("Step lengths") +
  theme_classic()

```

### Plot the distribution of next steps relative to the current location

```{r}

window_extent <- 2000

buffalo_all_steps %>% 
  filter(dt_hour < 24) %>%
  ggplot() +
  geom_point(aes(x = dx, y = dy), 
             fill = "black", colour = "black", 
             alpha = 0.1, size = 0.1) +
  scale_x_continuous("Next step in x direction (m)", limits = c(-window_extent, window_extent)) +
  scale_y_continuous("Next step in y direction (m)", limits = c(-window_extent, window_extent)) +
  coord_equal() +
  ggtitle("Next steps relative to current location") +
  theme_classic()

```

## Assess the buffer distance

```{r}

res <- 25 # resolution of the raster layers in metres

# proportion of steps less than the buffer distance
# the distance should be a multiple of the resolution
distance <- 1875

# check if multiple of the resolution?
distance %% res == 0

# buffer distance (add the resolution of the cell/2 as we want a centre cell)
buffer <- distance + (res/2)

# calculate the number of cells in each axis
nxn_cells <- buffer*2/res
print(paste0("Number of cells in each axis: ", nxn_cells))

# calculate proportion
prop_steps <- buffalo_all_steps %>% filter(dx > -distance & 
                                             dx < distance &
                                             dy > -distance & 
                                             dy < distance) %>% 
  nrow() / buffalo_all_steps %>% nrow()

print(paste0("Proportion of steps less than ", distance, "m in x or y direction: ", round(prop_steps, 4)))

# maxmimum distance at corner
max_dist <- sqrt(buffer^2 + buffer^2)
print(paste0("Maximum distance at corner: ", round(max_dist, 2)))

```

## Save the data frame

Save all of the data, as well as the two clusters separately.

```{r}

# remove steps that fall outside of the local spatial extent
buffalo_all_steps <- buffalo_all_steps %>%
  # filter(x2_ - x1_ > -distance & 
  #          x2_ - x1_ < distance &
  #          y2_ - y1_ > -distance & 
  #          y2_ - y1_ < distance) %>% 
  drop_na(ta_)

bigedi_steps <- buffalo_all_steps %>% filter(cluster == 1)
gropyul_steps <- buffalo_all_steps %>% filter(cluster == 2)

# save the full data frame
buffalo_all_steps %>% write_csv("data/mimal_steps.csv")

# save the steps for each cluster
bigedi_steps %>% write_csv("data/mimal_bigedi_steps.csv")
gropyul_steps %>% write_csv("data/mimal_gropyul_steps.csv")

```

# Spatial covariates

## Create an AOI for downloading DEM

```{r}

# Function to create rectangle from coordinate bounds
create_rectangle_shapefile <- function(df, x_col = "x", y_col = "y", 
                                       buffer = 0.01, crs = 4326, 
                                       output_path = "rectangle") {
  
  # Calculate bounds with buffer
  min_x <- min(df[[x_col]], na.rm = TRUE) - buffer
  max_x <- max(df[[x_col]], na.rm = TRUE) + buffer
  min_y <- min(df[[y_col]], na.rm = TRUE) - buffer
  max_y <- max(df[[y_col]], na.rm = TRUE) + buffer
  
  # Create rectangle coordinates (clockwise from bottom-left)
  rectangle_coords <- matrix(c(
    min_x, min_y,  # bottom-left
    max_x, min_y,  # bottom-right
    max_x, max_y,  # top-right
    min_x, max_y,  # top-left
    min_x, min_y   # close the polygon
  ), ncol = 2, byrow = TRUE)
  
  # Create polygon geometry
  rectangle_poly <- st_polygon(list(rectangle_coords))
  
  # Create sf object
  rectangle_sf <- st_sf(
    id = 1,
    area = st_area(rectangle_poly),
    geometry = st_sfc(rectangle_poly, crs = crs)
  )
  
  rectangle_projected <- st_transform(rectangle_sf, crs = 4326) 
  
  # Export as shapefile
  st_write(rectangle_projected, paste0(output_path, ".shp"), 
           delete_dsn = TRUE, quiet = TRUE)
  
  # Also export as GeoJSON (alternative format)
  st_write(rectangle_projected, paste0(output_path, ".geojson"), 
           delete_dsn = TRUE, quiet = TRUE)
  
  # Print summary
  cat("Rectangle created with bounds:\n")
  cat("X range:", min_x, "to", max_x, "\n")
  cat("Y range:", min_y, "to", max_y, "\n")
  cat("Files saved:", paste0(output_path, ".shp"), "and", paste0(output_path, ".geojson"), "\n")
  
  return(rectangle_projected)
}

```

### Use function to create shapefile and geojson

```{r}

# Create the rectangle shapefile
mimal_extent <- create_rectangle_shapefile(
  df = buffalo_all_steps,
  x_col = "x1_",           # column name for easting/longitude
  y_col = "y1_",           # column name for northing/latitude
  buffer = 5000,           # buffer distance in coordinate units
  crs = 3112,              # coordinate reference system of GPS locations (will output in 4326)
  output_path = "mapping/mimal_extent" # without extension
)

# View the result
print(mimal_extent)
plot(st_geometry(mimal_extent), main = "Rectangle from Coordinate Bounds")

# Optional: Add your original points to the plot
buffalo_all_steps_sf <- st_as_sf(buffalo_all_steps, coords = c("x1_", "y1_"), crs = 3112)
buffalo_all_steps_sf_wgs84 <- st_transform(buffalo_all_steps_sf, crs = 4326) 
plot(st_geometry(mimal_extent), main = "Rectangle with Original Points")
plot(st_geometry(buffalo_all_steps_sf_wgs84), add = TRUE, col = "red")#, pch = 16)

```

## Check DEM and create slope and roughness layers

We downloaded the DEM from [Geoscience Australia](https://elevation.fsdf.org.au/), which is a 1 second resolution DEM for Australia. We will use this to create slope and roughness layers. We will .

```{r}

# read in the DEM
# dem <- terra::rast("mapping/cropped rasters/DEM_Smoothed_mimal.tif") # smoothed DEM
dem <- terra::rast("mapping/cropped rasters/DEM_H_mimal.tif") # hydro-enforced DEM
dem
plot(dem)

```

## Reproject to match the Sentinel-2 layers

```{r}

s2 <- terra::rast("mapping/cropped rasters/sentinel2/25m/mimal/S2_SR_masked_mimal_2023_08.tif")
plot(s2[[1]])

# reproject to the coordinates of the buffalo data
dem <- terra::project(dem, s2)

# check the projection
terra::plot(dem)
terra::points(buffalo_all_steps[, c("x1_", "y1_")], col = "red")

# save the reprojected DEM
terra::writeRaster(dem, "mapping/cropped rasters/DEM_mimal_reprojected.tif", overwrite = T)

```

## Create slope and roughness layers

There doesn't appear to be much difference in appearance between the slope and roughness layers, so we can use either one. The slope layer is the angle of the terrain, while the roughness layer is a measure of the variability in slope over a given area. We will create both layers and save them as separate raster files.

```{r}

# create slope layer
slope <- terra::terrain(dem, v = "slope", unit = "degrees", neighbors = 8)
plot(slope)

# create roughness layer
roughness <- terra::terrain(dem, v = "roughness", unit = "degrees", neighbors = 8)
plot(roughness)

# save the layers
terra::writeRaster(slope, "mapping/cropped rasters/slope_mimal.tif", overwrite = T)
terra::writeRaster(roughness, "mapping/cropped rasters/roughness_mimal.tif", overwrite = T)

```


# Crop layers for environmental similarity anlaysis

```{r}

bigedi_steps 
gropyul_steps 

```

## Extent from df function

```{r}

calculate_extent_from_df <- function(df, 
                                     x_col = "x",
                                     y_col = "y", 
                                     buffer = 1000){

  # Calculate bounds with buffer
  min_x <- min(df[[x_col]], na.rm = TRUE) - buffer
  max_x <- max(df[[x_col]], na.rm = TRUE) + buffer
  min_y <- min(df[[y_col]], na.rm = TRUE) - buffer
  max_y <- max(df[[y_col]], na.rm = TRUE) + buffer
  
  extent <- terra::ext(min_x, max_x, min_y, max_y)
  
  return(extent)
  
}

```

## Create extent objects

```{r}

bigedi_extent <- calculate_extent_from_df(bigedi_steps, xcol = "x1_", y_col = "y1_", buffer = 2500)
gropyul_extent <- calculate_extent_from_df(gropyul_steps, xcol = "x1_", y_col = "y1_", buffer = 2500)

```

## Crop layers

```{r}

# load the october S2 data
s2 <- terra::rast("mapping/cropped rasters/sentinel2/25m/mimal/S2_SR_masked_mimal_2023_10.tif")

# crop the rasters
s2_bigedi <- terra::crop(s2, bigedi_extent)
dem_bigedi <- terra::crop(dem, bigedi_extent)
slope_bigedi <- terra::crop(slope, bigedi_extent)

# plot the rasters
plot(s2_bigedi[[1]])
terra::points(bigedi_steps[, c("x1_", "y1_")], col = "red")
plot(dem_bigedi)
plot(slope_bigedi)

# terra::writeRaster(s2_bigedi, "mapping/env_similarity/S2_mimal_2023_10_bigedi.tif", overwrite = T)

```






# Creating some layers for social dynamics social plot

```{r}

single_step <- buffalo_all_steps %>% slice(30000) %>% select(x1_, y1_)

```

## Plot the single location

```{r}

terra::plot(dem)
terra::points(single_step, col = "red")

```

## Crop out with a buffer

```{r}

buffer_dist <- 2000  # buffer distance in map units

# Create extent and crop
ext <- ext(single_step$x1_ - buffer_dist, 
           single_step$x1_ + buffer_dist, 
           single_step$y1_ - buffer_dist, 
           single_step$y1_ + buffer_dist)

```

### Crop the rasters

```{r}

# crop layers
cropped_dem <- crop(dem, ext)
cropped_slope <- crop(slope, ext)
cropped_s2_2 <- crop(s2[[2]], ext)
cropped_s2_3 <- crop(s2[[3]], ext)
cropped_s2_4 <- crop(s2[[4]], ext)

# normalise the layers from 0 to 1
cropped_dem <- (cropped_dem - min(cropped_dem[])) / (max(cropped_dem[]) - min(cropped_dem[]))
cropped_slope <- (cropped_slope - min(cropped_slope[])) / (max(cropped_slope[]) - min(cropped_slope[]))
cropped_s2_2 <- (cropped_s2_2 - min(cropped_s2_2[])) / (max(cropped_s2_2[]) - min(cropped_s2_2[]))
cropped_s2_3 <- (cropped_s2_3 - min(cropped_s2_3[])) / (max(cropped_s2_3[]) - min(cropped_s2_3[]))
cropped_s2_4 <- (cropped_s2_4 - min(cropped_s2_4[])) / (max(cropped_s2_4[]) - min(cropped_s2_4[]))

# plot cropped layers
plot(cropped_dem, main = "Cropped DEM")
plot(cropped_slope, main = "Cropped Slope")
plot(cropped_s2_2, main = "Cropped Sentinel-2 Band 2")
plot(cropped_s2_3, main = "Cropped Sentinel-2 Band 3")
plot(cropped_s2_4, main = "Cropped Sentinel-2 Band 4")

```

## Combine the layers

```{r}

habitat_selection <- 5 * cropped_dem - 1.25 * cropped_dem^2 +
  -5 * cropped_slope +
  cropped_s2_2 +
  cropped_s2_3 +
  5 * cropped_s2_4


plot(habitat_selection, main = "Habitat selection")

png("outputs/social_subnetwork/habitat_selection.png", width = 1500, height = 1000, res = 250)
plot(habitat_selection, main = "Habitat selection")
dev.off()

```

## Create a movement layers

```{r}

## Create distance and bearing layers for the movement probability
image_dim <- 160
pixel_size <- 25
center <- image_dim %/% 2

# Create matrices of indices
x <- matrix(rep(0:(image_dim - 1), image_dim), nrow = image_dim, byrow = TRUE)
y <- matrix(rep(0:(image_dim - 1), each = image_dim), nrow = image_dim, byrow = TRUE)

# Compute the distance layer
distance_layer <- sqrt((pixel_size * (x - center))^2 + (pixel_size * (y - center))^2)

# Change the center cell to the average distance from the center to all locations within the pixel
distance_layer[center + 1, center + 1] <- 0.3826 * pixel_size

# Compute the bearing layer
bearing_layer <- atan2(center - y, x - center)

# Convert the distance and bearing matrices to raster layers
distance_layer <- rast(distance_layer)
bearing_layer <- rast(bearing_layer)

plot(distance_layer, main = "Distance Layer")
plot(bearing_layer, main = "Bearing Layer")

```

## Calculate Gamma density

```{r}

# Gamma distribution parameters
shape = 2
scale = 500

# Get the distance values from the distance layer
distance_values <- terra::values(distance_layer)
# Use the distance_layer raster as a template
gamma_step_dist_2D <- distance_layer
# Calculate the gamma density values at the distances from centre
gamma_step_dist_2D[] <- dgamma(distance_values, 
                                shape = shape, 
                                scale = scale)

# Plot the gamma density layer
plot(gamma_step_dist_2D, main = "Gamma Density Layer")

```

## von Mises distribution

```{r}

# von Mises density function
dvonmises <- function(x, mu, kappa) {
  return(exp(kappa * cos(x - mu)) / (2 * pi * besselI(kappa, nu = 0)))
}

bearing_values <- terra::values(bearing_layer)
terra::plot(bearing_layer, main = "Bearing")
vonMises_dist_2D <- bearing_layer

vonMises_dist_2D[] <- dvonmises(bearing_values, mu = 2.5, kappa = 0.5)

terra::plot(vonMises_dist_2D, main = "von Mises Density Layer")

```

## Combine the layers

```{r}

movement_kernel <- gamma_step_dist_2D * vonMises_dist_2D
plot(movement_kernel, main = "Movement Kernel")

png("outputs/social_subnetwork/movement.png", width = 1500, height = 1000, res = 250)
plot(movement_kernel, main = "Movement Kernel")
dev.off()

crs(movement_kernel) <- crs(habitat_selection)
ext(habitat_selection) <- ext(movement_kernel)

movement_habitat <- habitat_selection * movement_kernel
plot(movement_habitat, main = "Movement Habitat Selection")

```

## Social dynamics layer

```{r}

## Create distance and bearing layers for the movement probability
animal_location_x <- 50
animal_location_y <- 70
# Compute the distance layer
social_layer_1 <- sqrt((pixel_size * (x - animal_location_x))^2 + (pixel_size * (y - animal_location_y))^2)
# Convert the social layer to raster layers
social_layer_1 <- rast(social_layer_1)
plot(social_layer_1, main = "Social layer 1")

animal_location_x <- 110
animal_location_y <- 50
social_layer_2 <- sqrt((pixel_size * (x - animal_location_x))^2 + (pixel_size * (y - animal_location_y))^2)
social_layer_2 <- rast(social_layer_2)
plot(social_layer_2, main = "Social layer 2")

animal_location_x <- 70
animal_location_y <- 130
social_layer_3 <- sqrt((pixel_size * (x - animal_location_x))^2 + (pixel_size * (y - animal_location_y))^2)
social_layer_3 <- rast(social_layer_3)
plot(social_layer_3, main = "Social layer 3")

```

## Combine the social layers

```{r}

social_layer <- (-5 * social_layer_1) - 1*social_layer_2 + 7*social_layer_3
social_layer <- (social_layer - min(social_layer[])) / (max(social_layer[]) - min(social_layer[]))
plot(social_layer, main = "Social Layer")

```

```{r}

# Define your setup
locations <- data.frame(
  x = c(50, 100, 110, 50, 60),  # x coordinates
  y = c(50, 110, 75, 120, 100),  # y coordinates
  sd = c(15, 20, 15, 10, 5)     # standard deviations for each location
)

# Create empty raster
r <- movement_habitat

# Get raster coordinates
coords <- xyFromCell(r, 1:ncell(r))

# Initialize values
values <- numeric(ncell(r))

# Add each normal distribution
for(i in 1:nrow(locations)) {
  dist <- sqrt((coords[,1] - locations$x[i])^2 + (coords[,2] - locations$y[i])^2)
  values <- values + exp(-0.5 * (dist/locations$sd[i])^2)
}

# Assign values to raster
values(r) <- values

plot(r, main = "Social layer")

png("outputs/social_subnetwork/social_layer.png", width = 1500, height = 1000, res = 250)
plot(r, main = "Social layer")
dev.off()

```

```{r}

# Define your setup
locations <- data.frame(
  x = c(50, 100, 110, 50, 60),  # x coordinates
  y = c(50, 110, 75, 120, 100),  # y coordinates
  sd = c(2, 2, 2, 2, 2)     # standard deviations for each location
)

# Create empty raster
r_locs <- movement_habitat

# Get raster coordinates
coords <- xyFromCell(r_locs, 1:ncell(r))

# Initialize values
values <- numeric(ncell(r_locs))

# Add each normal distribution
for(i in 1:nrow(locations)) {
  dist <- sqrt((coords[,1] - locations$x[i])^2 + (coords[,2] - locations$y[i])^2)
  values <- values + exp(-0.5 * (dist/locations$sd[i])^2)
}

# Assign values to raster
values(r_locs) <- values

plot(r_locs, main = "Animal locations")

png("outputs/social_subnetwork/animal_locations.png", width = 1500, height = 1000, res = 250)
plot(r_locs, main = "Animal locations")
dev.off()

```

# Combine all processes

```{r}

# Combine all the layers into a single raster stack
combined_layers <- habitat_selection * movement_kernel * r
plot(combined_layers, main = "Next-step probability")

png("outputs/social_subnetwork/next_step.png", width = 1500, height = 1000, res = 250)
plot(combined_layers, main = "Next-step probability")
dev.off()

```

