---
title: "deepSSF Data Prep"
author: "Scott Forrest"
date: "`r Sys.Date()`"
execute: 
  cache: false
bibliography: references.bib
toc: true
number-sections: false
format: 
  html:
    # self-contained: true
    code-fold: show
    code-tools: true
    df-print: paged
    code-line-numbers: true
    code-overflow: scroll
    fig-format: png
    fig-dpi: 300
  pdf:
    geometry: 
      - top=30mm
      - left=30mm
editor:
  source
abstract: |
  Here we prepare the data for fitting a deepSSF model. We load the GPS data and tidy it, create a trajectory object containing a series of steps and read in the environmental layers. For each observed step we crop out local subsets of the environmental layers, centred on the step's location. For every step we will then have the surrounding environmental information stored as a stack of small rasters, with the 'target' of what we're trying to predict (the actual location of the next step) as a raster layer as well, with all cells being 0 except the observed next location, which is 1. These, along with temporal covariates and the previous bearing, will be the input data for the deepSSF model.
---

## Loading packages

```{r}
#| warning=FALSE

library(tidyverse)
packages <- c("amt", "sf", "terra", "beepr", "tictoc")
walk(packages, require, character.only = T)

```

## Import data

```{r}

buffalo <- read_csv("data/SpaceCows/APN_processed.csv")

# the TS column has mixed formatting for the date time, 
# so we'll convert it using the individual columns for year, month, day, time
buffalo <- buffalo %>% mutate(
  datetime = lubridate::ymd_hms(paste(year, month, day, time)),
  .before = TS
)

sum(is.na(buffalo$datetime))

```

## Tidy data

```{r}

# arrange by time and exclude any duplicate timestamps (within the same individual's data)
buffalo <- buffalo %>%  
  group_by(device_id) %>% 
  arrange(datetime, .by_group = T) %>% 
  distinct(datetime, .keep_all = T) %>%
  arrange(device_id) %>% 
  mutate(id = device_id, 
         .before = datetime) %>% 
  ungroup()

# keep only the relevant columns
buffalo_clean <- buffalo[, c("id", "datetime", "Longitude", "Latitude")]

# rename the columns
colnames(buffalo_clean) <- c("id", "time", "lon", "lat")

# convert the time to the correct timezone
attr(buffalo_clean$time, "tzone") <- "Australia/Queensland"
head(buffalo_clean)

# check timezone
tz(buffalo_clean$time)

# create a vector of the unique ids for indexing later
buffalo_ids <- unique(buffalo_clean$id)

min(buffalo_clean$lon)
max(buffalo_clean$lon)
min(buffalo_clean$lat)
max(buffalo_clean$lat)

```

## Setup trajectory

Use the `amt` package to create a trajectory object from the cleaned data. 

```{r}

buffalo_all <- buffalo_clean %>% mk_track(id = id,
                                          lon,
                                          lat, 
                                          time, 
                                          all_cols = T,
                                          crs = 4326) %>% 
  # Transformation to GDA94 / Geoscience Australia Lambert (https://epsg.io/3112)
  transform_coords(crs_to = 3112, crs_from = 4326) %>% arrange(id) 

# Number of IDs and locations
buffalo_all %>%
  summarise(n_ids = n_distinct(id),
            n_locations = n())

```

## Plot the data coloured by time

```{r}

buffalo_all %>%
  ggplot(aes(x = x_, y = y_, colour = id)) +
  geom_point(alpha = 0.5, size = 0.1) + 
  coord_fixed() +
  scale_x_continuous("Easting (m)") +
  scale_y_continuous("Northing (m)") +
  scale_colour_viridis_d() +
  theme_classic() +
  theme(legend.position = "none") 

ggsave("outputs/data_prep/buffalo_APN_map.png", 
       width = 150, height = 150, units = "mm",  dpi = 1000)

```

## Plot a timeline of each individual

```{r}

buffalo_all %>% 
  ggplot(aes(x = t_, y = as.factor(id), colour = as.factor(id))) +
  geom_point(alpha = 0.5, size = 0.01) + 
  scale_x_datetime("Date") +
  scale_y_discrete("Individual ID") +
  scale_colour_viridis_d() +
  theme_classic() +
  theme(legend.position = "none",
        axis.text.y = element_blank()) 

ggsave("outputs/data_prep/buffalo_APN_timeline.png", 
       width = 150, height = 150, units = "mm",  dpi = 600)

```

## Number of location per day

```{r}

buffalo_daily_locations <- buffalo_all %>% 
  mutate(date = lubridate::date(t_)) %>% 
  dplyr::group_by(date) %>% 
  summarise(
    n_locations = n(),
    n_individuals = n_distinct(id)
)

buffalo_daily_locations %>% 
  ggplot() + 
    geom_col(aes(x = date, y = n_locations), 
             fill = "skyblue") + 
    geom_col(aes(x = date, y = n_individuals), 
             fill = "orange") + 
    scale_x_date("Date") +
    scale_y_continuous("Number of IDs/Locations") +
    scale_colour_viridis_d() +
    theme_classic() +
    theme(legend.position = "none") 

ggsave("outputs/data_prep/buffalo_APN_daily_locations.png",
       width = 150, height = 80, units = "mm",  dpi = 600)

```

# Generating the data to fit a deepSSF model

## Create a steps object to check the step lengths

```{r}

# nest the data by individual
buffalo_all_nested <- buffalo_all %>% arrange(id) %>% nest(data = -"id")

buffalo_all_nested_steps <- buffalo_all_nested %>%
  mutate(steps = map(data, function(x)
    x %>% 
      # track_resample(rate = hours(1), tolerance = minutes(10)) %>%
      steps()))

# unnest the data after creating 'steps' objects
buffalo_all_steps <- buffalo_all_nested_steps %>% 
  amt::select(id, steps) %>% 
  amt::unnest(cols = steps)

head(buffalo_all_steps)

```

## Prepare and export data for single layer saving (Python script)

We also have another script that saves all local layers as individual npy arrays, which works better when there are many steps per id (to prevent large tif files), and can make things a bit easier for training models with many ids.

```{r}

buffalo_all_steps <- buffalo_all_steps %>% mutate(
  
  # difference in the x and y directions
  dx = x2_ - x1_,
  dy = y2_ - y1_,
  
  dt_minute = as.numeric(round(difftime(t2_, t1_, units = "mins"), 0)),
  dt_hour = as.numeric(round(difftime(t2_, t1_, units = "hours"), 2)),
  
  # add the minutes as a decimal to the hour, and the hour as a decimal to the day
  hour_t1 = round(lubridate::hour(t1_) + lubridate::minute(t1_)/60,2),
  yday_t1 = round(lubridate::yday(t1_) + lubridate::hour(t1_)/24,2),
  
  hour_t1_sin1 = sin(2*pi*hour_t1/24),
  hour_t1_cos1 = cos(2*pi*hour_t1/24),
  hour_t1_sin2 = sin(4*pi*hour_t1/24),
  hour_t1_cos2 = cos(4*pi*hour_t1/24),
  
  yday_t1_sin1 = sin(2*pi*yday_t1/365.25),
  yday_t1_cos1 = cos(2*pi*yday_t1/365.25),
  yday_t1_sin2 = sin(4*pi*yday_t1/365.25),
  yday_t1_cos2 = cos(4*pi*yday_t1/365.25),
  
  # add the minutes as a decimal to the hour, and the hour as a decimal to the day
  hour_t2 = round(lubridate::hour(t2_) + lubridate::minute(t2_)/60,2),
  yday_t2 = round(lubridate::yday(t2_) + lubridate::hour(t2_)/24,2),
  
  bearing = direction_p,
  bearing_sin = sin(bearing),
  bearing_cos = cos(bearing)
  
)

buffalo_all_steps

```

### Plot step lengths

```{r}

buffalo_all_steps %>% 
  filter(dt_hour < 24) %>%
  filter(sl_ < 5000) %>%
  ggplot() +
  geom_density(aes(x = sl_), #, bins = 50
                 fill = "skyblue", colour = "skyblue", alpha = 0.75) +
  scale_x_continuous("Step length (m)") + #, limits = c(-25, 1250)
  # scale_x_log10("Step length (m)") + #, limits = c(-25, 1250)
  scale_y_continuous("Density") +
  ggtitle("Step lengths") +
  theme_classic()

```

### Step lengths vs step duration

```{r}

buffalo_all_steps %>% 
  filter(dt_hour < 24) %>%
  ggplot() +
  geom_point(aes(x = dt_hour, y = sl_), 
                 fill = "black", colour = "black", 
             alpha = 0.05, size = 0.1) +
  scale_x_continuous("Time between GPS locations (hours)") + #, limits = c(-25, 1250)
  scale_y_continuous("Step length (m)") +
  # scale_x_log10("Time between GPS locations (hours)") + #, limits = c(-25, 1250)
  scale_y_log10("Step length (m)") +
  ggtitle("Step lengths") +
  theme_classic()

```

### Plot the distribution of next steps relative to the current location

```{r}

window_extent <- 2000

buffalo_all_steps %>% 
  filter(dt_hour < 24) %>%
  ggplot() +
  geom_point(aes(x = dx, y = dy), 
             fill = "black", colour = "black", 
             alpha = 0.1, size = 0.1) +
  scale_x_continuous("Next step in x direction (m)", limits = c(-window_extent, window_extent)) +
  scale_y_continuous("Next step in y direction (m)", limits = c(-window_extent, window_extent)) +
  coord_equal() +
  ggtitle("Next steps relative to current location") +
  theme_classic()

```

## Set the buffer distance

```{r}

res <- 25 # resolution of the raster layers in metres

# proportion of steps less than the buffer distance
# the distance should be a multiple of the resolution
distance <- 1875

# check if multiple of the resolution?
distance %% res == 0

# buffer distance (add the resolution of the cell/2 as we want a centre cell)
buffer <- distance + (res/2)

# calculate the number of cells in each axis
nxn_cells <- buffer*2/res
print(paste0("Number of cells in each axis: ", nxn_cells))

# calculate proportion
prop_steps <- buffalo_all_steps %>% filter(x2_ - x1_ > -distance & 
                                             x2_ - x1_ < distance &
                                             y2_ - y1_ > -distance & 
                                             y2_ - y1_ < distance) %>% 
  nrow() / buffalo_all_steps %>% nrow()

print(paste0("Proportion of steps less than ", distance, "m in x or y direction: ", round(prop_steps, 4)))

# maxmimum distance at corner
max_dist <- sqrt(buffer^2 + buffer^2)
print(paste0("Maximum distance at corner: ", round(max_dist, 2)))

```

## Save the dataframe

```{r}

# remove steps that fall outside of the local spatial extent
buffalo_all_steps <- buffalo_all_steps %>%
  # filter(x2_ - x1_ > -distance & 
  #          x2_ - x1_ < distance &
  #          y2_ - y1_ > -distance & 
  #          y2_ - y1_ < distance) %>% 
  drop_na(ta_)

buffalo_all_steps

buffalo_all_steps %>% write_csv("data/APN_steps.csv")

```

# Spatial covariates

## Create an AOI for downloading DEM

```{r}

# Function to create rectangle from coordinate bounds
create_rectangle_shapefile <- function(df, x_col = "x", y_col = "y", 
                                       buffer = 0.01, crs = 4326, 
                                       output_path = "rectangle") {
  
  # Calculate bounds with buffer
  min_x <- min(df[[x_col]], na.rm = TRUE) - buffer
  max_x <- max(df[[x_col]], na.rm = TRUE) + buffer
  min_y <- min(df[[y_col]], na.rm = TRUE) - buffer
  max_y <- max(df[[y_col]], na.rm = TRUE) + buffer
  
  # Create rectangle coordinates (clockwise from bottom-left)
  rectangle_coords <- matrix(c(
    min_x, min_y,  # bottom-left
    max_x, min_y,  # bottom-right
    max_x, max_y,  # top-right
    min_x, max_y,  # top-left
    min_x, min_y   # close the polygon
  ), ncol = 2, byrow = TRUE)
  
  # Create polygon geometry
  rectangle_poly <- st_polygon(list(rectangle_coords))
  
  # Create sf object
  rectangle_sf <- st_sf(
    id = 1,
    area = st_area(rectangle_poly),
    geometry = st_sfc(rectangle_poly, crs = crs)
  )
  
  rectangle_projected <- st_transform(rectangle_sf, crs = 4326) 
  
  # Export as shapefile
  st_write(rectangle_projected, paste0(output_path, ".shp"), 
           delete_dsn = TRUE, quiet = TRUE)
  
  # Also export as GeoJSON (alternative format)
  st_write(rectangle_projected, paste0(output_path, ".geojson"), 
           delete_dsn = TRUE, quiet = TRUE)
  
  # Print summary
  cat("Rectangle created with bounds:\n")
  cat("X range:", min_x, "to", max_x, "\n")
  cat("Y range:", min_y, "to", max_y, "\n")
  cat("Files saved:", paste0(output_path, ".shp"), "and", paste0(output_path, ".geojson"), "\n")
  
  return(rectangle_projected)
}

```

### Use function to create shapefile and geojson

```{r}

# Create the rectangle shapefile
mimal_extent <- create_rectangle_shapefile(
  df = buffalo_all_steps,
  x_col = "x1_",           # column name for easting/longitude
  y_col = "y1_",           # column name for northing/latitude
  buffer = 5000,           # buffer distance in coordinate units
  crs = 3112,              # coordinate reference system of GPS locations (will output in 4326)
  output_path = "mapping/APN_extent" # without extension
)

# View the result
print(mimal_extent)
plot(st_geometry(mimal_extent), main = "Rectangle from Coordinate Bounds")

# Optional: Add your original points to the plot
buffalo_all_steps_sf <- st_as_sf(buffalo_all_steps, coords = c("x1_", "y1_"), crs = 3112)
buffalo_all_steps_sf_wgs84 <- st_transform(buffalo_all_steps_sf, crs = 4326) 
plot(st_geometry(mimal_extent), main = "Rectangle with Original Points")
plot(st_geometry(buffalo_all_steps_sf_wgs84), add = TRUE, col = "red")#, pch = 16)

```

## Check DEM and create slope and roughness layers

We downloaded the DEM from [Geoscience Australia](https://elevation.fsdf.org.au/), which is a 1 second resolution DEM for Australia. We will use this to create slope and roughness layers. We will .

```{r}

# read in the DEM
dem <- terra::rast("mapping/cropped rasters/DEM_H_APN.tif") # hydro-enforced DEM
dem
plot(dem)

```

## Reproject to match the Sentinel-2 layers

```{r}

s2 <- terra::rast("mapping/cropped rasters/sentinel2/25m/APN/S2_SR_masked_APN_2023_08.tif")
terra::plot(s2[[1]])

# reproject to the coordinates of the buffalo data
dem <- terra::project(dem, s2)

# check the projection
terra::plot(dem)
terra::points(buffalo_all_steps[, c("x1_", "y1_")], col = "red")

# save the reprojected DEM
terra::writeRaster(dem, "mapping/cropped rasters/DEM_APN_reprojected.tif", overwrite = T)

```

## Create slope and roughness layers

There doesn't appear to be much difference in appearance between the slope and roughness layers, so we can use either one. The slope layer is the angle of the terrain, while the roughness layer is a measure of the variability in slope over a given area. We will create both layers and save them as separate raster files.

```{r}

# create slope layer
slope <- terra::terrain(dem, v = "slope", unit = "degrees", neighbors = 8)
plot(slope)

# create roughness layer
roughness <- terra::terrain(dem, v = "roughness", unit = "degrees", neighbors = 8)
plot(roughness)

# save the layers
terra::writeRaster(slope, "mapping/cropped rasters/slope_APN.tif", overwrite = T)
terra::writeRaster(roughness, "mapping/cropped rasters/roughness_APN.tif", overwrite = T)

```