---
title: "deepSSF Data Prep"
author: "Scott Forrest"
date: "`r Sys.Date()`"
execute: 
  cache: false
bibliography: references.bib
toc: true
number-sections: false
format: 
  html:
    # self-contained: true
    code-fold: show
    code-tools: true
    df-print: paged
    code-line-numbers: true
    code-overflow: scroll
    fig-format: png
    fig-dpi: 300
  pdf:
    geometry: 
      - top=30mm
      - left=30mm
editor:
  source
abstract: |
  Here we prepare the data for fitting a deepSSF model. We load the GPS data and tidy it, create a trajectory object containing a series of steps and read in the environmental layers. For each observed step we crop out local subsets of the environmental layers, centred on the step's location. For every step we will then have the surrounding environmental information stored as a stack of small rasters, with the 'target' of what we're trying to predict (the actual location of the next step) as a raster layer as well, with all cells being 0 except the observed next location, which is 1. These, along with temporal covariates and the previous bearing, will be the input data for the deepSSF model.
---

## Loading packages

```{r}
#| warning=FALSE

library(tidyverse)
packages <- c("amt", "sf", "terra", "beepr", "tictoc")
walk(packages, require, character.only = T)

```

## Import data

```{r}

buffalo <- read_csv("data/SpaceCows/Lakefield_processed.csv")

# the TS column has mixed formatting for the date time, 
# so we'll convert it using the individual columns for year, month, day, time
buffalo <- buffalo %>% mutate(
  datetime = lubridate::ymd_hms(paste(year, month, day, time)),
  .before = TS
)

sum(is.na(buffalo$datetime))

```

## Tidy data

```{r}

# arrange by time and exclude any duplicate timestamps (within the same individual's data)
buffalo <- buffalo %>%  
  group_by(device_id) %>% 
  arrange(datetime, .by_group = T) %>% 
  distinct(datetime, .keep_all = T) %>%
  arrange(device_id) %>% 
  mutate(id = device_id, 
         .before = datetime) %>% 
  ungroup()

# keep only the relevant columns
buffalo_clean <- buffalo[, c("id", "datetime", "Longitude", "Latitude")]

# rename the columns
colnames(buffalo_clean) <- c("id", "time", "lon", "lat")

# convert the time to the correct timezone
attr(buffalo_clean$time, "tzone") <- "Australia/Queensland"
head(buffalo_clean)

# check timezone
tz(buffalo_clean$time)

# create a vector of the unique ids for indexing later
buffalo_ids <- unique(buffalo_clean$id)

min(buffalo_clean$lon)
max(buffalo_clean$lon)
min(buffalo_clean$lat)
max(buffalo_clean$lat)

```

## Setup trajectory

Use the `amt` package to create a trajectory object from the cleaned data. 

```{r}

buffalo_all <- buffalo_clean %>% mk_track(id = id,
                                          lon,
                                          lat, 
                                          time, 
                                          all_cols = T,
                                          crs = 4326) %>% 
  # Transformation to GDA94 / Geoscience Australia Lambert (https://epsg.io/3112)
  transform_coords(crs_to = 3112, crs_from = 4326) %>% arrange(id) 

```

## Plot the data coloured by time

```{r}

buffalo_all %>%
  ggplot(aes(x = x_, y = y_, colour = id)) +
  geom_point(alpha = 0.5, size = 0.1) + 
  coord_fixed() +
  scale_x_continuous("Easting (m)") +
  scale_y_continuous("Northing (m)") +
  scale_colour_viridis_d() +
  theme_classic() +
  theme(legend.position = "none") 

ggsave("outputs/data_prep/buffalo_lakefield_map.png", 
       width = 150, height = 150, units = "mm",  dpi = 1000)

```

## Plot a timeline of each individual

```{r}

buffalo_all %>% 
  ggplot(aes(x = t_, y = as.factor(id), colour = as.factor(id))) +
  geom_point(alpha = 0.5, size = 0.01) + 
  scale_x_datetime("Date") +
  scale_y_discrete("Individual ID") +
  scale_colour_viridis_d() +
  theme_classic() +
  theme(legend.position = "none",
        axis.text.y = element_blank()) 

ggsave("outputs/data_prep/buffalo_lakefield_timeline.png", 
       width = 150, height = 150, units = "mm",  dpi = 600)

```

## Number of location per day

```{r}

buffalo_daily_locations <- buffalo_all %>% 
  mutate(date = lubridate::date(t_)) %>% 
  dplyr::group_by(date) %>% 
  summarise(
    n_locations = n(),
    n_individuals = n_distinct(id)
)

buffalo_daily_locations %>% 
  ggplot() + 
    geom_col(aes(x = date, y = n_locations), 
             fill = "skyblue") + 
    geom_col(aes(x = date, y = n_individuals), 
             fill = "orange") + 
    scale_x_date("Date") +
    scale_y_continuous("Number of IDs/Locations") +
    scale_colour_viridis_d() +
    theme_classic() +
    theme(legend.position = "none") 

ggsave("outputs/data_prep/buffalo_lakefield_daily_locations.png",
       width = 150, height = 80, units = "mm",  dpi = 600)

```

# Generating the data to fit a deepSSF model

## Create a steps object to check the step lengths

```{r}

# nest the data by individual
buffalo_all_nested <- buffalo_all %>% arrange(id) %>% nest(data = -"id")

buffalo_all_nested_steps <- buffalo_all_nested %>%
  mutate(steps = map(data, function(x)
    x %>% 
      # track_resample(rate = hours(1), tolerance = minutes(10)) %>%
      steps()))

# unnest the data after creating 'steps' objects
buffalo_all_steps <- buffalo_all_nested_steps %>% 
  amt::select(id, steps) %>% 
  amt::unnest(cols = steps)

head(buffalo_all_steps)

```

## Prepare and export data for single layer saving (Python script)

We also have another script that saves all local layers as individual npy arrays, which works better when there are many steps per id (to prevent large tif files), and can make things a bit easier for training models with many ids.

```{r}

buffalo_all_steps <- buffalo_all_steps %>% mutate(
  
  dt_minute = as.numeric(round(difftime(t2_, t1_, units = "mins"), 0)),
  dt_hour = as.numeric(round(difftime(t2_, t1_, units = "hours"), 2)),
  
  # add the minutes as a decimal to the hour, and the hour as a decimal to the day
  hour_t1 = round(lubridate::hour(t1_) + lubridate::minute(t1_)/60,2),
  yday_t1 = round(lubridate::yday(t1_) + lubridate::hour(t1_)/24,2),
  
  hour_t1_sin1 = sin(2*pi*hour_t1/24),
  hour_t1_cos1 = cos(2*pi*hour_t1/24),
  hour_t1_sin2 = sin(4*pi*hour_t1/24),
  hour_t1_cos2 = cos(4*pi*hour_t1/24),
  
  yday_t1_sin1 = sin(2*pi*yday_t1/365.25),
  yday_t1_cos1 = cos(2*pi*yday_t1/365.25),
  yday_t1_sin2 = sin(4*pi*yday_t1/365.25),
  yday_t1_cos2 = cos(4*pi*yday_t1/365.25),
  
  # add the minutes as a decimal to the hour, and the hour as a decimal to the day
  hour_t2 = round(lubridate::hour(t2_) + lubridate::minute(t2_)/60,2),
  yday_t2 = round(lubridate::yday(t2_) + lubridate::hour(t2_)/24,2),
  
  bearing = direction_p,
  bearing_sin = sin(bearing),
  bearing_cos = cos(bearing)
  
)

buffalo_all_steps

```

### Plot step lengths

```{r}

buffalo_all_steps %>% 
  filter(dt_hour < 24) %>%
  filter(sl_ < 5000) %>%
  ggplot() +
  geom_density(aes(x = sl_), #, bins = 50
                 fill = "skyblue", colour = "skyblue", alpha = 0.75) +
  scale_x_continuous("Step length (m)") + #, limits = c(-25, 1250)
  # scale_x_log10("Step length (m)") + #, limits = c(-25, 1250)
  scale_y_continuous("Density") +
  ggtitle("Step lengths") +
  theme_classic()

```

### Step lengths vs step duration

```{r}

buffalo_all_steps %>% 
  filter(dt_hour < 24) %>%
  ggplot() +
  geom_point(aes(x = dt_hour, y = sl_), 
                 fill = "black", colour = "black", 
             alpha = 0.05, size = 0.1) +
  scale_x_continuous("Time between GPS locations (hours)") + #, limits = c(-25, 1250)
  scale_y_continuous("Step length (m)") +
  # scale_x_log10("Time between GPS locations (hours)") + #, limits = c(-25, 1250)
  scale_y_log10("Step length (m)") +
  ggtitle("Step lengths") +
  theme_classic()

```

## Set the buffer distance

```{r}

res <- 25 # resolution of the raster layers in metres

# proportion of steps less than the buffer distance
# the distance should be a multiple of the resolution
distance <- 1875

# check if multiple of the resolution?
distance %% res == 0

# buffer distance (add the resolution of the cell/2 as we want a centre cell)
buffer <- distance + (res/2)

# calculate proportion
prop_steps <- buffalo_all_steps %>% filter(x2_ - x1_ > -distance & 
                                             x2_ - x1_ < distance &
                                             y2_ - y1_ > -distance & 
                                             y2_ - y1_ < distance) %>% 
  nrow() / buffalo_all_steps %>% nrow()

print(paste0("Proportion of steps less than ", distance, "m in x or y direction: ", round(prop_steps, 4)))

# maxmimum distance at corner
max_dist <- sqrt(buffer^2 + buffer^2)
print(paste0("Maximum distance at corner: ", round(max_dist, 2)))

```



```{r}

# remove steps that fall outside of the local spatial extent
buffalo_all_steps_save <- buffalo_all_steps_save %>%
  filter(x2_ - x1_ > -distance & 
           x2_ - x1_ < distance &
           y2_ - y1_ > -distance & 
           y2_ - y1_ < distance) %>% 
  drop_na(ta_)

buffalo_all_steps_save

# buffalo_all_steps_save %>% write_csv("data/buffalo_mimal_all_steps.csv")

```

In our case over 96% of the steps were less than a distance of 1262.5m in either the x or y direction, which would mean that including 101 x 101 cells (which would be 2525 m x 2525 m) would include at least 96% of the steps. As the distance is further to the corner (but still less than the buffer distance in the x and y directions), there will be step lengths longer than the buffer distance (up to 1785 m).

## Set up the spatial extent of the local covariates

We calculate the number of cells in each axis, and set the lag between locations. Typically this will just be 1, as we want the next-step to be the target. However, it may be possible to improve model training by using different lags and including the time difference between locations as a covariate. I expect that this would help to predict across different time scales and not be restricted to the time scale that the data was collected at.

```{r}

# calculate the number of cells in each axis
nxn_cells <- buffer*2/res
print(paste0("Number of cells in each axis: ", nxn_cells))

# hourly lag - to set larger time differences between locations
hourly_lag <- 1

```

# Loop over each individual and save the local rasters

This is the main function of the script. It loops over each individual and calculates step level information, such as the location and time of the next step (x2, y2, t2), step length, turning angle, and different time components. 

It also determines where to crop the local layers by subtracting and adding the buffer distance from the location of each step. It then crops out the local layers for each step and saves them as a list entry into the data frame. 

It then creates a raster stack (with the number of layers equal to the number of steps) for each covariate and saves it as a tif.

For each individual, this will therefore create a csv containing the step level information, and a tif for each covariate containing the local layers for each step. 

It takes quite a while to run, around 1 hour per individual in our case, so we will illustrate the function with 10 steps per individual. Uncomment the line specified in the loop to use the full dataset.


```{r}

# for(i in 1:length(buffalo_ids)) {
for(i in 1:1) { # select just the first ID

  buffalo_data <- buffalo_all %>% filter(id == buffalo_ids[i])
  # all data for that individual
  buffalo_data <- buffalo_data %>% arrange(t_)
  
  # to use a subset of the data for that individual for testing
  # COMMENT THIS LINE OUT TO USE THE FULL DATASET
  buffalo_data <- buffalo_data %>% arrange(t_) |>  slice(1:10)

  n_samples <- nrow(buffalo_data)
  
  tic()
  
  buffalo_data_covs <- buffalo_data %>% mutate(
    
    x1_ = x_,
    y1_ = y_,
    x2_ = lead(x1_, n = hourly_lag, default = NA),
    y2_ = lead(y1_, n = hourly_lag, default = NA),
    x2_cent = x2_ - x1_,
    y2_cent = y2_ - y1_,
    t2_ = lead(t_, n = hourly_lag, default = NA),
    t_diff = round(difftime(t2_, t_, units = "hours"),0),
    
    # add the minutes as a decimal to the hour, and the hour as a decimal to the day
    hour_t1 = round(lubridate::hour(t_) + lubridate::minute(t_)/60,2),
    yday_t1 = round(lubridate::yday(t_) + lubridate::hour(t_)/24,2),
    
    hour_t1_sin1 = sin(2*pi*hour_t1/24),
    hour_t1_cos1 = cos(2*pi*hour_t1/24),
    hour_t1_sin2 = sin(4*pi*hour_t1/24),
    hour_t1_cos2 = cos(4*pi*hour_t1/24),
    hour_t1_sin3 = sin(6*pi*hour_t1/24),
    hour_t1_cos3 = cos(6*pi*hour_t1/24),
    
    yday_t1_sin1 = sin(2*pi*yday_t1/365.25),
    yday_t1_cos1 = cos(2*pi*yday_t1/365.25),
    yday_t1_sin2 = sin(4*pi*yday_t1/365.25),
    yday_t1_cos2 = cos(4*pi*yday_t1/365.25),
    yday_t1_sin3 = sin(6*pi*yday_t1/365.25),
    yday_t1_cos3 = cos(6*pi*yday_t1/365.25),
    
    # add the minutes as a decimal to the hour, and the hour as a decimal to the day
    hour_t2 = round(lubridate::hour(t2_) + lubridate::minute(t2_)/60,2),
    hour_t2_sin = sin(2*pi*hour_t2/24),
    hour_t2_cos = cos(2*pi*hour_t2/24),
    
    yday_t2 = round(lubridate::yday(t2_) + lubridate::hour(t2_)/24,2),
    yday_t2_sin = sin(2*pi*yday_t2/365.25),
    yday_t2_cos = cos(2*pi*yday_t2/365.25),
    
    sl = c(sqrt(diff(y_)^2 + diff(x_)^2), NA),
    log_sl = log(sl),
    bearing = c(atan2(diff(y_), diff(x_)), NA),
    bearing_sin = sin(bearing),
    bearing_cos = cos(bearing),
    ta = c(NA, ifelse(
      diff(bearing) > pi, diff(bearing)-(2*pi), ifelse(
        diff(bearing) < -pi, diff(bearing)+(2*pi), diff(bearing)))),
    cos_ta = cos(ta),
      
    # extent for cropping the spatial covariates
    x_min = x_ - buffer,
    x_max = x_ + buffer,
    y_min = y_ - buffer,
    y_max = y_ + buffer,
    
  
    # crop out and store the local covariates centered on the animal's location 
    # with an extent set in the previous chunk
  ) %>% rowwise() %>% mutate(

    # the extent needs to be the same for saving the stack of rasters
    extent_00centre = list(ext(x_min - x_, x_max - x_, y_min - y_, y_max - y_)),

    # NDVI
    ndvi_index = which.min(abs(difftime(t_, terra::time(ndvi_projected)))),
    ndvi_cent = list({
      ndvi_cent = crop(ndvi_projected[[ndvi_index]], ext(x_min, x_max, y_min, y_max))
      ext(ndvi_cent) <- extent_00centre
      ndvi_cent
      }),

    # herbaceous vegetation
    veg_herby_cent = list({
      veg_herby_cent = crop(veg_herby, ext(x_min, x_max, y_min, y_max))
      ext(veg_herby_cent) <- extent_00centre
      veg_herby_cent
      }),

    # canopy cover
    canopy_cover_cent = list({
      canopy_cover_cent = crop(canopy_cover, ext(x_min, x_max, y_min, y_max))
      ext(canopy_cover_cent) <- extent_00centre
      canopy_cover_cent
      }),

    # slope
    slope_cent = list({
      slope_cent <- crop(slope, ext(x_min, x_max, y_min, y_max))
      ext(slope_cent) <- extent_00centre
      slope_cent
      }),
    
    pres_template = list({
      pres_template <- crop(slope, ext(x_min, x_max, y_min, y_max))
      pres_template
    }),
    
    
    points_vect_cent_x1 = list(terra::vect(cbind(x_, y_), type = "points", crs = "EPSG:3112")),
    
    pres_cent_x1 = list({
      pres_cent_x1 <- rasterize(points_vect_cent_x1, pres_template, background=0)
      # ext(pres_cent_x1) <- extent_00centre
      # ext(pres_cent_x1) <- ext(x_min, x_max, y_min, y_max) # should already have this extent
      pres_cent_x1
    }),
    
    # Recentering method
    # Can pull locations into different cells
    # rasterised location of the next step - centred on (0,0)
    points_vect_cent_x2_recentred = list(terra::vect(cbind(x2_ - x_, y2_ - y_), type = "points", crs = "EPSG:3112")),
    pres_cent_x2_recentred = list({
      pres_cent_x2_recentred <- rasterize(points_vect_cent_x2_recentred, ndvi_cent, background=0)
      # ext(pres_cent_x2) <- extent_00centre # should already have this extent
      # ext(pres_cent_x2_recentred) <- ext(pres_cent_x1) # to set the natural extent
      pres_cent_x2_recentred
      }),
    
    # Strict within cell method
    points_vect_cent_x2_within_cell = list(terra::vect(cbind(x2_, y2_), type = "points", crs = "EPSG:3112")),
    pres_cent_x2_within_cell = list({
      pres_cent_x2_within_cell <- rasterize(points_vect_cent_x2_within_cell, pres_template, background=0)
      ext(pres_cent_x2_within_cell) <- extent_00centre
      pres_cent_x2_within_cell
      })
    
  ) %>% ungroup() # to remove the 'rowwise' class


  toc()

  # remove steps that fall outside of the local spatial extent
  buffalo_data_covs <- buffalo_data_covs %>%
    filter(x2_cent > -buffer & x2_cent < buffer & y2_cent > -buffer & y2_cent < buffer) %>%
    drop_na(ta)

  # remove the columns that are no longer needed
  buffalo_data_df <- buffalo_data_covs %>%
  dplyr::select(
    -extent_00centre,
    -ndvi_cent,
    -veg_herby_cent,
    -canopy_cover_cent,
    -slope_cent,
    -pres_template,
    -points_vect_cent_x1,
    -pres_cent_x1,
    -points_vect_cent_x2_recentred,
    # -pres_cent_x2_recentred,
    -points_vect_cent_x2_within_cell,
    -pres_cent_x2_within_cell)

  # # save the data
  # write_csv(buffalo_data_df, paste0("buffalo_local_data_id/buffalo_temporal_cont_", buffalo_ids[i],
  #                                   "_data_df_lag_", hourly_lag, "hr_n", n_samples, ".csv"))
  # 
  # # saving the raster objects
  # rast(buffalo_data_covs$ndvi_cent) %>%
  #   writeRaster(paste0("buffalo_local_layers_id/buffalo_", buffalo_ids[i], "_ndvi_cent",
  #                      nxn_cells, "x", nxn_cells, "_lag_", hourly_lag, "hr_n", n_samples, ".tif"),
  #               overwrite = T)
  # 
  # rast(buffalo_data_covs$veg_herby_cent) %>%
  #   writeRaster(paste0("buffalo_local_layers_id/buffalo_", buffalo_ids[i], "_herby_cent",
  #                      nxn_cells, "x", nxn_cells, "_lag_", hourly_lag, "hr_n", n_samples, ".tif"),
  #               overwrite = T)
  # 
  # rast(buffalo_data_covs$canopy_cover_cent) %>%
  #   writeRaster(paste0("buffalo_local_layers_id/buffalo_", buffalo_ids[i], "_canopy_cent",
  #                      nxn_cells, "x", nxn_cells, "_lag_", hourly_lag, "hr_n", n_samples, ".tif"),
  #               overwrite = T)
  # 
  # rast(buffalo_data_covs$slope_cent) %>%
  #   writeRaster(paste0("buffalo_local_layers_id/buffalo_", buffalo_ids[i], "_slope_cent",
  #                      nxn_cells, "x", nxn_cells, "_lag_", hourly_lag, "hr_n", n_samples, ".tif"),
  #               overwrite = T)
  
  
  # rast(buffalo_data_covs$pres_cent_x2_recentred) %>%
  #   writeRaster(paste0("buffalo_local_layers_id/recentred_buffalo_", buffalo_ids[i], "_pres_cent",
  #                      nxn_cells, "x", nxn_cells, "_lag_", hourly_lag, "hr_n", n_samples, ".tif"),
  #               overwrite = T)

  # rast(buffalo_data_covs$pres_cent_x2_within_cell) %>%
  #   writeRaster(paste0("buffalo_local_layers_id/within_cell_buffalo_", buffalo_ids[i], "_pres_cent",
  #                      nxn_cells, "x", nxn_cells, "_lag_", hourly_lag, "hr_n", n_samples, ".tif"),
  #               overwrite = T)
  
}

```

## Check the outputs

```{r}

head(buffalo_data_df)
head(buffalo_data_covs)
which(is.na(buffalo_data_covs$ta))

```

# Check which cell is occupied by the next step (the target)

```{r}

n_samples <- 8

print("Current step")

# x1
for(i in 1:n_samples) {
  print(rowColFromCell(buffalo_data_covs$pres_cent_x1[[i]], which(buffalo_data_covs$pres_cent_x1[[i]][] == 1)))
}

print("Next step (target) recentered")

# x2 recentered
for(i in 1:n_samples) {
  print(rowColFromCell(buffalo_data_covs$pres_cent_x2_recentred[[i]], which(buffalo_data_covs$pres_cent_x2_recentred[[i]][] == 1)))
}

print("Next step (target) within cell")

# x2 centered
for(i in 1:n_samples) {
  print(rowColFromCell(buffalo_data_covs$pres_cent_x2_within_cell[[i]], which(buffalo_data_covs$pres_cent_x2_within_cell[[i]][] == 1)))
}

```

## Checking the distribution of next-steps after the discretisation

```{r}

original_target <- rast("buffalo_local_layers_id/buffalo_2005_pres_cent101x101_lag_1hr_n10297.tif")
fixed_target <- rast("buffalo_local_layers_id/fixed_buffalo_2005_pres_cent101x101_lag_1hr_n10297.tif")

# fixed_target <- rast("buffalo_local_layers_id/within_cell_buffalo_2005_pres_cent101x101_lag_1hr_n100.tif")

```

```{r}

n_samples <- nlyr(original_target)
# n_samples <- 100

print("Uncorrected target")

x_vals <- rep(NA, n_samples)
y_vals <- rep(NA, n_samples)
distance <- rep(NA, n_samples)

# x2 uncentered
tic()
for(i in 1:n_samples) {
  
  # plot(original_target[[i]])
  
  # print(rowColFromCell(original_target[[i]], which(original_target[[i]][] == 1)))
  # print(xyFromCell(original_target[[i]], which(original_target[[i]][] == 1)))
  
  xy <- xyFromCell(original_target[[i]], which(original_target[[i]][] == 1))
  
  x_vals[i] <- xy[1]
  y_vals[i] <- xy[2]
  
  distance[i] <- sqrt((x_vals[i])^2 + (y_vals[i])^2)
  
}
toc()

```

### Do the same for the corrected target

```{r}

n_samples <- nlyr(fixed_target)
# n_samples <- 97

print("Corrected target")

x_vals_corr <- rep(NA, n_samples)
y_vals_corr <- rep(NA, n_samples)
distance_corr <- rep(NA, n_samples)

# x2 centered
tic()
for(i in 1:n_samples) {
  
  # plot(original_target[[i]])
  # print(rowColfromCell(original_target[[i]], which(original_target[[i]][] == 1)))
  # print(xyFromCell(original_target[[i]], which(original_target[[i]][] == 1)))
  xy <- xyFromCell(fixed_target[[i]], which(fixed_target[[i]][] == 1))
  
  x_vals_corr[i] <- xy[1]
  y_vals_corr[i] <- xy[2]
  
  distance_corr[i] <- sqrt((x_vals_corr[i])^2 + (y_vals_corr[i])^2)
  
}
toc()

distance[1:100] == distance_corr[1:100]

```


### Check the discretised distribution against the observed data

```{r}

buffalo_data <- read_csv("buffalo_local_data_id/buffalo_temporal_cont_2005_data_df_lag_1hr_n10297.csv")
buffalo_data$discrete_sl_uncorr <- distance
buffalo_data$discrete_sl_corr <- distance_corr

```

## Checking the observed distribution of small steps 

```{r}

buffalo_data %>% 
  filter(discrete_sl_corr == 0) %>% 
  summarise(mean_sl = mean(sl, na.rm = T),
            median_sl = median(sl, na.rm = T))

buffalo_data %>% 
  filter(discrete_sl_uncorr == 0) %>% 
  ggplot() +
  geom_density(aes(x = sl), 
                 fill = "skyblue", colour = "skyblue", alpha = 0.25) +
  scale_x_continuous("Step length (m)") +
  scale_y_continuous("Density") +
  ggtitle("Step lengths") +
  theme_classic()

```


```{r}

buffalo_data <- buffalo_data %>% mutate(
  discrete_sl_uncorr = ifelse(discrete_sl_uncorr == 0, 10, discrete_sl_uncorr),
  discrete_sl_uncorr_jitter = discrete_sl_uncorr + rnorm(nrow(buffalo_data), 0, 6.5),
  # discrete_sl_uncorr_jitter = discrete_sl_uncorr + runif(nrow(buffalo_data), 0, 25),
  
  discrete_sl_corr = ifelse(discrete_sl_corr == 0, 10, discrete_sl_corr),
  discrete_sl_corr_jitter = discrete_sl_corr + rnorm(nrow(buffalo_data), 0, 6.5),
  # discrete_sl_corr_jitter = discrete_sl_corr + runif(nrow(buffalo_data), 0, 25)
)

```


```{r}

# fixed centre cell adjustment
buffalo_data %>% 
  filter(sl < 2000) %>%
  ggplot() +
  geom_density(aes(x = sl), 
                 fill = "skyblue", colour = "skyblue", alpha = 0.25) +
  geom_density(aes(x = discrete_sl_uncorr),
                 fill = "purple", colour = "purple", alpha = 0.25) +
  geom_density(aes(x = discrete_sl_corr),
                 fill = "orange", colour = "orange", alpha = 0.25) +
  scale_x_continuous("Step length (m)") +
  # scale_x_log10("Step length (m)") + 
  scale_y_continuous("Density") +
  ggtitle("Step lengths") +
  theme_classic()

# jitter for every cell
buffalo_data %>% 
  filter(discrete_sl_uncorr < 2000) %>%
  ggplot() +
  geom_density(aes(x = sl), 
                 fill = "skyblue", colour = "skyblue", alpha = 0.25) +
  geom_density(aes(x = discrete_sl_uncorr_jitter), 
                 fill = "purple", colour = "purple", alpha = 0.25) +
  geom_density(aes(x = discrete_sl_corr_jitter), 
                 fill = "orange", colour = "orange", alpha = 0.25) +
  scale_x_continuous("Step length (m)") +
  # scale_x_log10("Step length (m)") + 
  scale_y_continuous("Density") +
  ggtitle("Step lengths") +
  theme_classic()

# fixed centre cell adjustment
buffalo_data %>% 
  filter(discrete_sl_uncorr < 2000) %>%
  ggplot() +
  geom_density(aes(x = sl), 
                 fill = "skyblue", colour = "skyblue", alpha = 0.25) +
  geom_density(aes(x = discrete_sl_uncorr), 
                 fill = "purple", colour = "purple", alpha = 0.25) +
  geom_density(aes(x = discrete_sl_corr), 
                 fill = "orange", colour = "orange", alpha = 0.25) +
  # scale_x_continuous("Step length (m)") + 
  scale_x_log10("Step length (m)") +
  scale_y_continuous("Density") +
  ggtitle("Step lengths") +
  theme_classic()

# jitter for every cell
buffalo_data %>% 
  filter(discrete_sl_uncorr < 2000) %>%
  ggplot() +
  geom_density(aes(x = sl), 
                 fill = "skyblue", colour = "skyblue", alpha = 0.25) +
  geom_density(aes(x = discrete_sl_uncorr_jitter), 
                 fill = "purple", colour = "purple", alpha = 0.25) +
  geom_density(aes(x = discrete_sl_corr_jitter), 
                 fill = "orange", colour = "orange", alpha = 0.25) +
  # scale_x_continuous("Step length (m)") + 
  scale_x_log10("Step length (m)") +
  scale_y_continuous("Density") +
  ggtitle("Step lengths") +
  theme_classic()

```



## Plot a subset of the local covariates

```{r}

n_plots <- 5

# NDVI
walk(buffalo_data_covs$ndvi_cent[1:n_plots], terra::plot)

# Herbaceous vegetation
walk(buffalo_data_covs$veg_herby_cent[1:n_plots], terra::plot)

# Canopy cover
walk(buffalo_data_covs$canopy_cover_cent[1:n_plots], terra::plot)

# Slope
walk(buffalo_data_covs$slope_cent[1:n_plots], terra::plot)

# Target (location of the next step)
walk(buffalo_data_covs$pres_cent_x1[1:n_plots], terra::plot)

# Target (location of the next step)
walk(buffalo_data_covs$pres_cent_x2_within_cell[1:n_plots], terra::plot)

```


```{r}

dummy_extent <- terra::ext(buffalo_data_covs$pres_cent_x1[[1]])
(dummy_extent[2] - dummy_extent[1])/2 + dummy_extent[1]
(dummy_extent[4] - dummy_extent[3])/2 + dummy_extent[3]

```

The recentred approach (uses the difference between the (x1, y1) and (x2, y2) locations).

```{r}

for(i in 1:nrow(buffalo_data_covs)) {
  
  combo_recentred_df <- as.data.frame(buffalo_data_covs$pres_cent_x1[[i]] + buffalo_data_covs$pres_cent_x2_recentred[[i]], xy = TRUE)
  
  plot <- ggplot() +
    geom_raster(data = combo_recentred_df, aes(x, y, fill = last)) +
    geom_point(data = buffalo_data_covs[i, ], aes(x = x_, y = y_), colour = "black", size = 2) +
    geom_point(data = buffalo_data_covs[i, ], aes(x = x2_, y = y2_), colour = "red", size = 2) +
    coord_fixed() +
    theme_bw() +
    theme(legend.position = "none") +
    ggtitle(paste0("Step ", i))
  
  print(plot)
  
}

```

```{r}

for(i in 1:nrow(buffalo_data_covs)) {
  
  combo_within_cell_df <- as.data.frame(buffalo_data_covs$pres_cent_x1[[i]] + buffalo_data_covs$pres_cent_x2_within_cell[[i]], xy = TRUE)
  
  plot <- ggplot() +
    geom_raster(data = combo_within_cell_df, aes(x, y, fill = last)) +
    geom_point(data = buffalo_data_covs[i, ], aes(x = x_, y = y_), colour = "black", size = 2) +
    geom_point(data = buffalo_data_covs[i, ], aes(x = x2_, y = y2_), colour = "red", size = 2) +
    coord_fixed() +
    theme_bw() +
    theme(legend.position = "none") +
    ggtitle(paste0("Step ", i))
  
  print(plot)
  
}

```


## To save the object with all of the local covariates

```{r}

# saveRDS(buffalo_data_covs, paste0("buffalo_data_id_covs_", Sys.Date(), ".rds"))

```

## To save some plots

```{r}

# for(i in 1:n_plots) {
#   png(filename = paste0("ndvi_cent_", i, ".png"),
#       width = 150, height = 150, units = "mm", res = 600)
#   terra::plot(buffalo_data_covs$ndvi_cent[[i]])
#   dev.off()
# }
# 
# # for the pres layers
# for(i in 1:n_plots) {
#   
#   # change the 0 values to NA
#   layer <- buffalo_data_covs$pres_cent[[i]]
#   layer[layer == 0] <- 0.5
#   
#   ndvi_layer <- buffalo_data_covs$ndvi_cent[[i]]
#   new_layer <- layer*ndvi_layer
#   
#   # save the plot
#   png(filename = paste0("pres_cent_", i, ".png"),
#       width = 150, height = 150, units = "mm", res = 600)
#   terra::plot(new_layer)
#   dev.off()
# }
# 
# 
# # other plots
# 
# n_plots <- 1
# 
# for(i in 1:n_plots) {
#   png(filename = paste0("veg_herby_cent_", i, ".png"),
#       width = 150, height = 150, units = "mm", res = 600)
#   terra::plot(buffalo_data_covs$veg_herby_cent[[i]])
#   dev.off()
# }
# 
# for(i in 1:n_plots) {
#   png(filename = paste0("canopy_cover_cent_", i, ".png"),
#       width = 150, height = 150, units = "mm", res = 600)
#   terra::plot(buffalo_data_covs$canopy_cover_cent[[i]])
#   dev.off()
# }
# 
# for(i in 1:n_plots) {
#   png(filename = paste0("slope_cent_", i, ".png"),
#       width = 150, height = 150, units = "mm", res = 600)
#   terra::plot(buffalo_data_covs$slope_cent[[i]])
#   dev.off()
# }

```

## Plot the step lengths and turning angles

```{r}

# plot step lengths
buffalo_data_covs %>% ggplot() +
  geom_density(aes(x = sl), 
                 fill = "skyblue", colour = "skyblue", alpha = 0.75) +
  scale_x_continuous("Step length (m)") + 
  scale_y_continuous("Density") +
  ggtitle("Step lengths") +
  theme_classic()

# ggsave("outputs/step_length.png", height = 60, width = 120, units = "mm", dpi = 600)


# plot turning angles
buffalo_data_covs %>% ggplot() +
  geom_density(aes(x = ta), 
                 fill = "skyblue", colour = "skyblue", alpha = 0.75) +
  scale_x_continuous("Turning angle (radians)", 
    breaks = c(-pi, -pi/2, 0, pi/2, pi),
    labels = c(expression(-pi), expression(-pi/2), "0", expression(pi/2), expression(pi))
  ) +
  scale_y_continuous("Density") +
  ggtitle("Turning angles") +
  theme_classic()

# ggsave("outputs/turning_angle.png", height = 60, width = 120, units = "mm", dpi = 600)  

```

